---
title: "01_Deconvolution_deconBenchmark"
author: "IMEV2"
date: "`r Sys.Date()`"
output: html_document
---
# 0 Environment
## 0.1 Packages
### 0.1.1 Installs
```{r echo=F, message=F, warning=FALSE}
# if (!requireNamespace("devtools", quietly = TRUE)) {
#    install.packages("devtools")
# }
# devtools::install_github("Danko-Lab/BayesPrism/BayesPrism")
# install.packages("pheatmap")
```

### 0.1.1 Library
```{r echo=F, message=F, warning=FALSE}
set.seed(123)
library(babelwhale)
library(DeconBenchmark)
library(tidyverse)
library(SingleCellExperiment)
library(openxlsx)
library(readxl)
library(ggplot2)
library(patchwork)
library(pheatmap)
library(dplyr)
```

## 0.2 Paths
```{r echo=F, message=FALSE, warning=FALSE}
fprefix <- "03_Deconvolution_CTVT_deconBenchmark"
wDir <- paste0("~/gitClones/01_Deconvolution/03_Deconvolution_CTVT/")
dataDir <- paste0(wDir, "/031_Data/")
resDir <- paste0(wDir, "/033_Results/", fprefix, "/")
plotDir <- paste0(wDir, "/034_Plots/", fprefix, "/")
scriptDir <- paste0(wDir, "/032_Scripts/", fprefix, "/")

source(paste0(scriptDir, "03_Deconvolution_CTVT_deconBenchmark_metrics.R"))
```
## 0.3 Functions
### 0.3.1 Subsample from reference 
Use sample+cell type threshold.
```{r}
subsampleCells <- function(df, 
                            cell_type_col, 
                            total_cells = 10000, 
                            threshold = 50, 
                            seed = 123) {
  set.seed(seed)
  
  # Count total cells per type
  cell_counts <- table(df[[cell_type_col]])
  cell_types <- names(cell_counts)
  
  # Initialize sampling plan
  min_counts <- if (!is.null(threshold)) {
    pmin(cell_counts, pmax(threshold, 0))
  } else {
    rep(0, length(cell_counts))
  }
  
  names(min_counts) <- cell_types
  
  # Ensure we don't oversample beyond what's available
  min_counts <- pmin(min_counts, cell_counts)
  
  # Calculate remaining cells to sample after ensuring threshold
  guaranteed_cells <- sum(min_counts)
  
  remaining_cells <- total_cells - guaranteed_cells
  
  if (remaining_cells < 0) {
    stop("Thresholds exceed total desired cell count.")
  }
  
  # Adjust the remaining proportions based on cell types that exceed their thresholds
  eligible_types <- cell_types[cell_counts > min_counts]
  eligible_counts <- cell_counts[eligible_types] - min_counts[eligible_types]
  prop_remaining <- eligible_counts / sum(eligible_counts)
  additional_cells <- round(prop_remaining * remaining_cells)
  
  # Adjust additional_cells to cover all min_counts
  add_cells_full <- setNames(rep(0, length(min_counts)), names(min_counts))
  add_cells_full[names(additional_cells)] <- additional_cells
  
  # Final sample sizes
  final_sample_sizes <- min_counts + add_cells_full

  # Adjust rounding to hit total_cells exactly
  diff <- total_cells - sum(final_sample_sizes)
  
  if (diff != 0) {
    idx <- order(-prop_remaining)
    additional_cells[eligible_types[idx[1:abs(diff)]]] <- additional_cells[eligible_types[idx[1:abs(diff)]]] + sign(diff)
  }
  
  # # Final number to sample per type
  # final_sample_sizes <- min_counts
  # final_sample_sizes[eligible_types] <- final_sample_sizes[eligible_types] + additional_cells
  
  # Sample the rows
  sampled_rows <- unlist(lapply(cell_types, function(ct) {
    n <- final_sample_sizes[ct]
    sample(which(df[[cell_type_col]] == ct), size = n, replace = FALSE)
  }))
  
  df[sampled_rows, , drop = FALSE]
}

```


# 1.0 Inputs
BayesPrism will take intersect of genes in reference and bulk samples.
Use unnormalized and untransformed count data.

## 1.1 Genes of interest
```{r}
adrianq1q4100 <- read.xlsx("~/gitClones/01_Deconvolution/03_Deconvolution_CTVT/033_Results/03_Deconvolution_CTVT_geneCorrelations/03_Deconvolution_CTVT_geneCorrelations_mergedFamWilcoxDegs_q1q4_100Genes.xlsx") 

goi <- unique(adrianq1q4100$Symbol)
idoi <- unique(adrianq1q4100$GeneID)

geneDict <- adrianq1q4100 %>%
  select(Symbol, GeneID) %>%
  # Take distinct as marker genes may overlap clusters. 
  distinct()
```

## 1.1 Reference
Reference - cellxgene dense matrix.
### 1.1.1 Matrix
```{r}
obj <- readRDS(paste0(dataDir, "SCE_concat_CTVT_includingAllGenes_logCorrecDone_vstCorrecDone_ClustDone2_ManAnot.rds"))

objNoBad <- obj[,obj$manual.log.anot.fine!="bad"]

rm(obj)
gc()


geneCorr_famDegs <- read.xlsx("~/gitClones/01_Deconvolution/03_Deconvolution_CTVT/033_Results/03_Deconvolution_CTVT_geneCorrelations/03_Deconvolution_CTVT_geneCorrelations_geneCorrelations_famDegs.xlsx")

snCountsFinal <- counts(objNoBad) %>%
  as.data.frame() %>%
  rownames_to_column("Symbol") %>%
  filter(Symbol %in% geneCorr_famDegs$Gene)

# Map Symbol to ID (in bulk)
rowData <- rowData(objNoBad) %>% 
  as.data.frame() %>%
  filter(Symbol %in% geneCorr_famDegs$Gene) %>%
  dplyr::select(Symbol, ID) 

### Each gene from snCountsFinal has a matching symbol. 

setdiff(snCountsFinal$Symbol, rowData$Symbol)

snCountsFinal <- snCountsFinal %>%
  left_join(rowData) %>%
  select(-Symbol) %>%
  column_to_rownames("ID")
  
snCountsFinal_colData <- colData(objNoBad) 

rm(objNoBad)
gc()
```

### 1.1.2 Subsample and filter genes
```{r}
snCountsFinal_colData$sampCtype <- paste0(snCountsFinal_colData$Sample,
                                          "_",
                                          snCountsFinal_colData$manual.log.anot.fine)

snCountsFinalSubSampled_colData <- subsampleCells(df = snCountsFinal_colData,
                                                    cell_type_col = "sampCtype",
                                                    total_cells = 30000, 
                                                    threshold = 50, 
                                                    seed = 123)

snCountsFinalSubSampled <- snCountsFinal %>%
  t() %>% # CHECK THIS WORKS!
  as.data.frame() %>%
  rownames_to_column("cellId") %>%
  # Select the rows in the counts matrix that correspond to cells from subsampled colData
  filter(cellId %in% rownames(snCountsFinalSubSampled_colData)) %>%
  column_to_rownames("cellId") %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  # select gene id's (stored as columns in df) that are hvgs.
  rownames_to_column("GeneID") %>%
  filter(GeneID %in% idoi) %>%
  left_join(geneDict) %>%
  column_to_rownames("Symbol") %>%
  select(-GeneID) %>%
  as.matrix()
```

### 1.1.2 Cell type annotations
Cell type labels - character vector same length as # cells in reference.
Cell state labels - optional, finer granularity character vector. (i.e. myeloid 1 vs myeloid 2). 
Labels should have at least 20 to 50 cells. 
```{r}
snAnno <- snCountsFinalSubSampled_colData$manual.log.anot.fine

print(sort(table(snAnno)))

# Create the cell count matrix: rows = cell types, columns = samples
snCellCounts <- snCountsFinalSubSampled_colData %>%
  as.data.frame() %>%
  group_by(manual.log.anot.fine, 
           Sample) %>%
  summarise(cellCount = n()) %>%
  pivot_wider(names_from = Sample, 
              values_from = cellCount, 
              values_fill = 0) %>%
  column_to_rownames("manual.log.anot.fine")
```
### 1.1.3 Ground truth 
```{r}
snCellCountsSubSampled <-snCountsFinalSubSampled_colData %>%
  as.data.frame() %>%
  group_by(manual.log.anot.fine,
           Sample) %>%
  summarise(cellCount = n()) %>%
  pivot_wider(names_from = Sample,
              values_from = cellCount,
              values_fill = 0) %>%
  column_to_rownames("manual.log.anot.fine")

snCellCountsSubSampled[,"3302T1p"] <- snCellCountsSubSampled[,"3302T1b"]

# Convert to percentages (column-wise normalization)
snCellPercentagesSubSampled <- sweep(snCellCountsSubSampled,
                           2,
                           colSums(snCellCountsSubSampled),
                           FUN = "/") %>%
  round(3) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    cols = -sample,
    names_to = "manual.log.anot.fine",
    values_to = "cellProportion"
  ) %>%
  # If sample name ends in b_n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
  mutate(
    sampleType = "sn",
    sampleName = sample,
    sampleNameEdit = str_remove(sampleName, "_p$|_2$")
  ) %>%
  select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
  mutate("deconvBenchResType" = "groundTruth", # P for proportion
         "Method" = "NA",
         "Parameters" = "NA",
         "Reference" = "subSampled",
         "Seed"=123)
```


# 1.3 Bulk
### 1.2.1 Matrix
```{r}
bulkCountsFinal <- readRDS(paste0(dataDir, "01_Deconvolution_matchedBulkMatrixCTVT_bulkCountsFinal.rds")) %>%
  as.data.frame() %>%
  # select gene id's (stored as columns in df) that are hvgs.
  rownames_to_column("GeneID") %>%
  filter(GeneID %in% idoi) %>%
  left_join(geneDict) %>%
  column_to_rownames("Symbol") %>%
  select(-GeneID) %>%
  as.matrix()

```

### 1.2.2 Metadata
```{r}
bulkCountsFinal_metadata <- read.xlsx(paste0(dataDir, "01_Deconvolution_matchedBulkMatrixCTVT_bulkCountsFinal_metadata.xlsx"))
```

# 3 Loop reference based deconvolution methods

## 3.1 Reference-based
I have included the method behind each method in the comment, as well as whether the method worked for this person who tested them all (https://github.com/tinnlab/DeconBenchmark/issues/4)
```{r}
print(getSupportedMethods())

masterPropDf <- data.frame(sampleName = c(), 
                           sampleType = c(), 
                           manual.log.anot.fine = c(), 
                           cellProportion =c(), 
                           sampleNameEdit = c(), 
                           deconvBenchResType = c(), 
                           Method = c(), 
                           Parameters = c())


masterMetricDf <- data.frame(bulkCellEstimate = c(), 
                             bulkNucEstimate = c(), 
                             Method = c(), 
                             Parameters = c(), 
                             Metric = c())
```

### 3.0.1 Debugging notes
See 00_deconBenchmark for more info. 
```{r}
babelwhale::test_docker_installation(detailed = TRUE) # Docker application needs to be running!!
Sys.getenv("DOCKER_DEFAULT_PLATFORM") # Should return [1] "linux/amd64" on Sanger Mac's. Change in .Rprofile if not the case. 
```

## 3.1 Relative proportions

### 3.1.1 HVGs
#### 3.1.1.1 Set 1
##### 3.1.1.1.1 MuSiC - Runs, second time not running?? 
- Can max out memory so lowered m to 14G (also check Activity Monitor - do not have Zotero running at the same time.)
```{r, warning=F, error=T, message=F}
MuSiC <- runDeconvolution("MuSiC", 
                         bulk = bulkCountsFinal,  # 23 Samples
                         singleCellExpr = snCountsFinalSubSampled, 
                         singleCellLabels = snAnno,
                         dockerArgs = c("--cpus=8.0",
                                        "-m=14G", 
                                        "--memory-reservation=8G"))

# Pivot results to "long" to be able to merge with the ground truth proportions from sn reference. 
musicRes <- MuSiC$MuSiC$P %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    cols = -sample,
    names_to = "manual.log.anot.fine",
    values_to = "cellProportion"
  ) %>%
    # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
    mutate(
      sampleType = case_when(
        str_ends(sample, "_n") ~ "bn",
        TRUE ~ "bc"
      ),
      # If sample type is bn, remove b_n to save sample id only.
      sampleName = case_when(
        sampleType == "bn" ~ str_remove(sample, "_n$"),
        TRUE ~ sample
      ),
      sampleName = case_when(
        sampleName == "3302T1b_p" ~ "3302T1p",
        TRUE ~ sampleName
      ),
      # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
      sampleNameEdit = case_when(
        sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
        TRUE ~ sampleName
      ),
       # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 

    ) %>%
  select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
  # Method details
  mutate("deconvBenchResType" = "P", # P for proportion
         "Method" = "MuSiC", 
         "Parameters" = "Default", 
         "Reference" = "subSampled",
         "Seed"=123)

# Rbind proportion outputs
musicResGround <- rbind(musicRes, snCellPercentagesSubSampled)

masterPropDf <- rbind(masterPropDf, musicResGround)

# Compute and rbind metric outputs
musicResMetrics <- deconError(musicResGround, 
                              parameters = "Default",
                              method = "MuSiC", 
                              reference="subSampled",
                              seed=123)
# masterMetricDf <- rbind(masterMetricDf, musicResMetrics)
# 

# Cleanup
rm(musicResGround, musicRes, musicResMetrics, MuSiC)
gc()
```
##### 3.1.1.1.2 RNA-Sieve - Runs
```{r}
RNASieve <- runDeconvolution("RNA-Sieve",
                         bulk = bulkCountsFinal,  # 20 Samples
                         singleCellExpr = snCountsFinalSubSampled,
                         singleCellLabels = snAnno,
                         dockerArgs = c("--cpus=8.0",
                                        "-m=16G",
                                        "--memory-reservation=8G"))

rnasieveRes <-RNASieve$`RNA-Sieve`$P %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    cols = -sample,
    names_to = "manual.log.anot.fine",
    values_to = "cellProportion"
  ) %>%
    # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
    mutate(
      sampleType = case_when(
        str_ends(sample, "_n") ~ "bn",
        TRUE ~ "bc"
      ),
      # If sample type is bn, remove b_n to save sample id only.
      sampleName = case_when(
        sampleType == "bn" ~ str_remove(sample, "_n$"),
        TRUE ~ sample
      ),
      sampleName = case_when(
        sampleName == "3302T1b_p" ~ "3302T1p",
        TRUE ~ sampleName
      ),
      # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
      sampleNameEdit = case_when(
        sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
        TRUE ~ sampleName
      ),
       # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 

    ) %>%
  select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
  # Method details
  mutate("deconvBenchResType" = "P", # P for proportion
         "Method" = "RNA-Sieve", 
         "Parameters" = "Default", 
         "Reference" = "subSampled",
         "Seed"=123)

# Rbind proportion outputs to compute error metrics.
rnasieveResGround <- rbind(rnasieveRes, snCellPercentagesSubSampled)

# snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
masterPropDf <- rbind(masterPropDf, rnasieveRes)


# Compute and rbind metric outputs
rnasieveResMetrics <- deconError(rnasieveResGround,
                              parameters = "Default",
                              method = "RNA-Sieve",
                              reference="subSampled",
                              seed=123)
# Rbind metric measurements
masterMetricDf <- rbind(masterMetricDf, rnasieveResMetrics)


# Cleanup
rm(rnasieveResGround, rnasieveRes, rnasieveResMetrics, RNASieve)
gc()
```


##### 3.1.1.1.4 Adroit - Runs
```{r}
AdRoit <- runDeconvolution("AdRoit", 
                         bulk = bulkCountsFinal,  
                         singleCellExpr = snCountsFinalSubSampled, 
                         singleCellLabels = snAnno,
                         dockerArgs = c("--cpus=8.0",
                                        "-m=16G", 
                                        "--memory-reservation=8G"))

adroitResP <- AdRoit$AdRoit$P %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    cols = -sample,
    names_to = "manual.log.anot.fine",
    values_to = "cellProportion"
  ) %>%
    # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
    mutate(
      sampleType = case_when(
        str_ends(sample, "_n") ~ "bn",
        TRUE ~ "bc"
      ),
      # If sample type is bn, remove b_n to save sample id only.
      sampleName = case_when(
        sampleType == "bn" ~ str_remove(sample, "_n$"),
        TRUE ~ sample
      ),
      sampleName = case_when(
        sampleName == "3302T1b_p" ~ "3302T1p",
        TRUE ~ sampleName
      ),
      # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
      sampleNameEdit = case_when(
        sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
        TRUE ~ sampleName
      ),
       # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 

    ) %>%
  select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
  # Method details
  mutate("deconvBenchResType" = "P", # P for proportion
         "Method" = "AdRoit", 
         "Parameters" = "Default", 
         "Reference" = "subSampled",
         "Seed"=123)


# Rbind proportion outputs to compute error metrics.
adroitResGround <- rbind(adroitResP, snCellPercentagesSubSampled)

# snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
masterPropDf <- rbind(masterPropDf, adroitResP)


# Compute and rbind metric outputs
adroitResMetrics <- deconError(adroitResGround,
                              parameters = "Default",
                              method = "AdRoit",
                              reference="subSampled",
                              seed=123)
# Rbind metric measurements
masterMetricDf <- rbind(masterMetricDf, adroitResMetrics$overviewMetrics)


# Cleanup
rm(adroitResGround, adroitResP, adroitResMetrics, AdRoit)
gc()
```
##### 3.1.1.1.5 digitalDLSorter 
Package no longer available on CRAN, requires editing docker. 


#### 3.1.1.2 Set 2
Of note, requires defining a signature matrix.
TODO: Check what method is "under the hood" for this function. 
```{r}
methodsToRunSet2 <- c("DWLS", # WWCLS
                  "MIXTURE", # vWCLS # Worked
                  "FARDEEP", # CLS # Worked
                  "LinDeconSeq",# WWCLS # Worked
                  "CIBERSORT", # Is this CIBERSORTx? # vWCLS
                  "BayesPrism", # Bayesian
                  "DecOT", # Ensembl
                  "AutoGeneS") # CLS/SVR
```

##### 3.1.1.2.1 DWLS
<!-- Check messages: -->
<!-- No dimnames in `exprsArray`, assuming `fData` and `cData` are sorted according to `exprsArray` -->
<!-- Assuming data assay in position 1, with name Et is log-transformed. -->
```{r}
# # $DWLS
# # [1] "bulk"             "singleCellExpr"   "singleCellLabels"
DWLS <- runDeconvolution("DWLS",
                         bulk = bulkCountsFinal,  # 20 Samples
                         singleCellExpr = snCountsFinalSubSampled,
                         singleCellLabels = snAnno,
                         dockerArgs = c("--cpus=8.0",
                                        "-m=16G",
                                        "--memory-reservation=8G"))
# dwlsResP <- DWLS$DWLS$P %>%
#   as.data.frame() %>%
#   
#   as.data.frame() %>%
#   rownames_to_column("sample") %>%
#   pivot_longer(
#     cols = -sample,
#     names_to = "manual.log.anot.fine",
#     values_to = "cellProportion"
#   ) %>%
#     # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
#     mutate(
#       sampleType = case_when(
#         str_ends(sample, "_n") ~ "bn",
#         TRUE ~ "bc"
#       ),
#       # If sample type is bn, remove b_n to save sample id only.
#       sampleName = case_when(
#         sampleType == "bn" ~ str_remove(sample, "_n$"),
#         TRUE ~ sample
#       ),
#       sampleName = case_when(
#         sampleName == "3302T1b_p" ~ "3302T1p",
#         TRUE ~ sampleName
#       ),
#       # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
#       sampleNameEdit = case_when(
#         sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
#         TRUE ~ sampleName
#       ),
#        # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 
# 
#     ) %>%
#   select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
#   # Method details
#   mutate("deconvBenchResType" = "P", # P for proportion
#          "Method" = "DWLS", 
#          "Parameters" = "Default", 
#          "Reference" = "subSampled",
#          "Seed"=123)
# 
# # Rbind proportion outputs to compute error metrics.
# dwlsResGround <- rbind(dwlsResP, snCellPercentagesSubSampled)
# 
# # snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
# masterPropDf <- rbind(masterPropDf, dwlsResP)
# 
# 
# # Compute and rbind metric outputs
# # dwlsResMetrics <- deconError(dwlsResGround, 
# #                               parameters = "Default",
# #                               method = "DWLS", 
# #                               reference="subSampled",
# #                               seed=123)
# # Rbind metric measurements 
# # masterMetricDf <- rbind(masterMetricDf, dwlsResMetrics)
# 
# 
# # Cleanup
# rm(dwlsResGround, dwlsResP, dwlsResMetrics, DWLS)
# gc()

# dwlsResS <- DWLS$DWLS$S %>%
#   as.data.frame() %>%
#   mutate("deconvBenchResType" = "S",
#          "Method" = "DWLS")
```

##### 3.1.1.2.2 MIXTURE
###### 3.1.1.2.2.2 All cell signature 
Still subset on HVGs
```{r}
# $MIXTURE
# [1] "bulk"      "signature"

sig <- generateReference(snCountsFinalSubSampled,
                         snAnno,
                         type="signature")

MIXTURE_fullRef <- runDeconvolution("MIXTURE", 
                         bulk = bulkCountsFinal,  # 20 Samples
                         signature = sig$signature, # 195 genes creating reference from hvgs and all cells rather than subsample. This could be indicative that we need to loop over multiple seeds for the true best method.  
                         dockerArgs = c("--cpus=8.0",
                                        "-m=16G", 
                                        "--memory-reservation=8G"))

mixtureResP_fullRef <- MIXTURE_fullRef$MIXTURE$P %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    cols = -sample,
    names_to = "manual.log.anot.fine",
    values_to = "cellProportion"
  ) %>%
    # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
    mutate(
      sampleType = case_when(
        str_ends(sample, "_n") ~ "bn",
        TRUE ~ "bc"
      ),
      # If sample type is bn, remove b_n to save sample id only.
      sampleName = case_when(
        sampleType == "bn" ~ str_remove(sample, "_n$"),
        TRUE ~ sample
      ),
      sampleName = case_when(
        sampleName == "3302T1b_p" ~ "3302T1p",
        TRUE ~ sampleName
      ),
      # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
      sampleNameEdit = case_when(
        sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
        TRUE ~ sampleName
      ),
       # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 

    ) %>%
  select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
  # Method details
  mutate("deconvBenchResType" = "P", # P for proportion
         "Method" = "MIXTURE", 
         "Parameters" = "fullSignatureMatrix", 
         "Reference" = "subSampled",
         "Seed"=123)

# # Rbind proportion outputs to compute error metrics.
# mixtureResGround <- rbind(mixtureResP_fullRef, snCellPercentagesSubSampled)

# snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
masterPropDf <- rbind(masterPropDf, mixtureResP_fullRef)


# # Compute and rbind metric outputs
# mixtureResMetrics <- deconError(mixtureResGround, 
#                               parameters = "fullSignatureMatrix",
#                               method = "MIXTURE", 
#                               reference="subSampled",
#                               seed=123)
# # Rbind metric measurements 
# masterMetricDf <- rbind(masterMetricDf, mixtureResMetrics)


# Cleanup
rm(mixtureResGround, mixtureResP, mixtureResMetrics, MIXTURE_fullRef)
gc()


# mixtureResS_fullRef <- MIXTURE_fullRef$MIXTURE$S %>%
#   as.data.frame() %>%
#   mutate("deconvBenchResType" = "S",
#          "Method" = "MIXTURE_fullRef")
```
##### 3.1.1.2.3 FARDEEP 
###### 3.1.1.2.3.1 All cell signature
```{r}
# # $FARDEEP
# # [1] "bulk"      "signature" --> technically our subset of genes is "signature
# 
# FARDEEP_fullRef <- runDeconvolution("FARDEEP", 
#                          bulk = bulkCountsFinal,  # 20 Samples
#                          signature = reference_HVGs$signature, # 195 marker genes only!! 
#                          dockerArgs = c("--cpus=8.0",
#                                         "-m=16G", 
#                                         "--memory-reservation=8G"))
# 
# fardeepResP_fullRef <- FARDEEP_fullRef$FARDEEP$P %>%
#   as.data.frame() %>%
#   rownames_to_column("sample") %>%
#   pivot_longer(
#     cols = -sample,
#     names_to = "manual.log.anot.fine",
#     values_to = "cellProportion"
#   ) %>%
#     # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
#     mutate(
#       sampleType = case_when(
#         str_ends(sample, "_n") ~ "bn",
#         TRUE ~ "bc"
#       ),
#       # If sample type is bn, remove b_n to save sample id only.
#       sampleName = case_when(
#         sampleType == "bn" ~ str_remove(sample, "_n$"),
#         TRUE ~ sample
#       ),
#       sampleName = case_when(
#         sampleName == "3302T1b_p" ~ "3302T1p",
#         TRUE ~ sampleName
#       ),
#       # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
#       sampleNameEdit = case_when(
#         sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
#         TRUE ~ sampleName
#       ),
#        # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 
# 
#     ) %>%
#   select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
#   # Method details
#   mutate("deconvBenchResType" = "P", # P for proportion
#          "Method" = "FARDEEP", 
#          "Parameters" = "fullSignatureMatrix", 
#          "Reference" = "subSampled",
#          "Seed"=123)
# 
# # Rbind proportion outputs to compute error metrics.
# # fardeepResGround <- rbind(fardeepResP_fullRef , snCellPercentagesSubSampled)
# 
# # snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
# masterPropDf <- rbind(masterPropDf, fardeepResP_fullRef)
# 
# 
# # # Compute and rbind metric outputs
# # fardeepResMetrics <- deconError(fardeepResGround, 
# #                               parameters = "fullSignatureMatrix",
# #                               method = "FARDEEP", 
# #                               reference="subSampled",
# #                               seed=123)
# # # Rbind metric measurements 
# # masterMetricDf <- rbind(masterMetricDf, fardeepResMetrics)
# 
# 
# # Cleanup
# rm(fardeepResGround, fardeepResP_fullRef , fardeepResMetrics, FARDEEP_fullRef)
# gc()
# 
# # fardeepResS <- FARDEEP_fullRef$FARDEEP$S %>%
# #   as.data.frame() %>%
# #   mutate("deconvBenchResType" = "S",
# #          "Method" = "FARDEEP")
# 
# ## Add full sig results. 
```
##### 3.1.1.2.4 LinDeconSeq

###### 3.1.1.2.4.1 All cell signature
```{r}
# # $LinDeconSeq
# # [1] "bulk"      "signature"
# LinDeconSeq_fullRef <- runDeconvolution("LinDeconSeq", 
#                          bulk = bulkCountsFinal,  # 20 Samples
#                          signature = reference_HVGs$signature,
#                          dockerArgs = c("--cpus=8.0",
#                                         "-m=16G", 
#                                         "--memory-reservation=8G"))
# 
# lindeconseqResP_fullRef <- LinDeconSeq_fullRef$LinDeconSeq$P %>%
#   as.data.frame() %>%
#   rownames_to_column("sample") %>%
#   pivot_longer(
#     cols = -sample,
#     names_to = "manual.log.anot.fine",
#     values_to = "cellProportion"
#   ) %>%
#     # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
#     mutate(
#       sampleType = case_when(
#         str_ends(sample, "_n") ~ "bn",
#         TRUE ~ "bc"
#       ),
#       # If sample type is bn, remove b_n to save sample id only.
#       sampleName = case_when(
#         sampleType == "bn" ~ str_remove(sample, "_n$"),
#         TRUE ~ sample
#       ),
#       sampleName = case_when(
#         sampleName == "3302T1b_p" ~ "3302T1p",
#         TRUE ~ sampleName
#       ),
#       # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
#       sampleNameEdit = case_when(
#         sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
#         TRUE ~ sampleName
#       ),
#        # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 
# 
#     ) %>%
#   select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
#   # Method details
#   mutate("deconvBenchResType" = "P", # P for proportion
#          "Method" = "LinDeconSeq", 
#          "Parameters" = "fullSignatureMatrix", 
#          "Reference" = "subSampled",
#          "Seed"=123)
# 
# # Rbind proportion outputs to compute error metrics.
# lindeconseqResGround <- rbind(lindeconseqResP_fullRef, snCellPercentagesSubSampled)
# 
# # snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
# masterPropDf <- rbind(masterPropDf, lindeconseqResP_fullRef)
# 
# 
# # # Compute and rbind metric outputs
# # lindeconseqResMetrics <- deconError(lindeconseqResGround, 
# #                               parameters = "fullSignatureMatrix",
# #                               method = "LinDeconSeq", 
# #                               reference="subSampled",
# #                               seed=123)
# # # Rbind metric measurements 
# # masterMetricDf <- rbind(masterMetricDf, lindeconseqResMetrics)
# 
# 
# # Cleanup
# rm(lindeconseqResGround, lindeconseqResP, lindeconseqResMetrics, LinDeconSeq_fullRef)
# gc()
```

##### 3.1.1.2.5 CIBERSORT 
Docker repo doesn't seem to exist??
```{r}
# CIBERSORT <- runDeconvolution("CIBERSORT", 
#                          bulk = bulkCountsFinal,  # 20 Samples
#                          singleCellExpr = snCountsFinalSubSampled, 
#                          singleCellLabels = snAnno,
#                          dockerArgs = c("--cpus=8.0",
#                                         "-m=16G", 
#                                         "--memory-reservation=8G"))
```
##### 3.1.1.2.6 BayesPrism
Takes veeeery long. Especially compared to Bayes Prism custom script. 
Will run seperately.
```{r}
# $BayesPrism
# [1] "bulk"             "singleCellExpr"   "singleCellLabels"

bayesPrism <- runDeconvolution("BayesPrism",
                         bulk = bulkCountsFinal,  # 20 Samples
                         singleCellExpr = snCountsFinalSubSampled,
                         singleCellLabels = snAnno,
                         dockerArgs = c("--cpus=8.0",
                                        "-m=14G",
                                        "--memory-reservation=8G"))

bayesprismResP <- bayesPrism$BayesPrism$P %>%
  as.data.frame() %>%
  
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    cols = -sample,
    names_to = "manual.log.anot.fine",
    values_to = "cellProportion"
  ) %>%
    # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
    mutate(
      sampleType = case_when(
        str_ends(sample, "_n") ~ "bn",
        TRUE ~ "bc"
      ),
      # If sample type is bn, remove b_n to save sample id only.
      sampleName = case_when(
        sampleType == "bn" ~ str_remove(sample, "_n$"),
        TRUE ~ sample
      ),
      sampleName = case_when(
        sampleName == "3302T1b_p" ~ "3302T1p",
        TRUE ~ sampleName
      ),
      # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
      sampleNameEdit = case_when(
        sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
        TRUE ~ sampleName
      ),
       # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 

    ) %>%
  select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
  # Method details
  mutate("deconvBenchResType" = "P", # P for proportion
         "Method" = "BayesPrism", 
         "Parameters" = "Default", 
         "Reference" = "subSampled",
         "Seed"=123)

# Rbind proportion outputs to compute error metrics.
bayesprismResGround <- rbind(bayesprismResP, snCellPercentagesSubSampled)

# snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
masterPropDf <- rbind(masterPropDf, bayesprismResP)


# # Compute and rbind metric outputs
# bayesprismResMetrics <- deconError(bayesprismResGround, 
#                               parameters = "Default",
#                               method = "BayesPrism", 
#                               reference="subSampled",
#                               seed=123)
# # Rbind metric measurements 
# masterMetricDf <- rbind(masterMetricDf, bayesprismResMetrics)


# Cleanup
rm(bayesprismResGround, bayesprismResP, bayesprismResMetrics, AutoGeneS)
gc()

```

##### 3.1.1.2.7 DecOT
Add MATLAB License. I can't find the license file here...
```{r}
# # $DecOT
# # [1] "bulk"               "singleCellExpr"     "singleCellLabels"  
# # [4] "singleCellSubjects"
# DecOT <- runDeconvolution("DecOT", 
#                          bulk = bulkCountsFinal,  # 20 Samples
#                          singleCellExpr = snCountsFinalSubSampled, 
#                          singleCellLabels = snAnno,
#                          singleCellSubjects = snSample_subSample_relProp,
#                          matlabLicenseFile = "~/Library/Application Support/MathWorks/MATLAB/R2024b/thisMatlab.pem",
#                          dockerArgs = c("--cpus=8.0",
#                                         "-m=16G", 
#                                         "--memory-reservation=8G"))
```
##### 3.1.1.2.8 AutoGeneS

###### 3.1.1.2.8.1 All cell cell type expr
```{r}
AutoGeneS_full <- runDeconvolution("AutoGeneS", 
                         bulk = bulkCountsFinal,  # 20 Samples
                         cellTypeExpr = reference_HVGs$cellTypeExpr,
                         dockerArgs = c("--cpus=8.0",
                                        "-m=16G", 
                                        "--memory-reservation=8G"))

autogenesResP <- AutoGeneS_full$AutoGeneS$P %>%
  as.data.frame() %>%
  rownames_to_column("sample") %>%
  pivot_longer(
    cols = -sample,
    names_to = "manual.log.anot.fine",
    values_to = "cellProportion"
  ) %>%
    # If sample name ends in _n, type is bn (bulk nuc). Otherwise, it is bc (bulk cell)
    mutate(
      sampleType = case_when(
        str_ends(sample, "_n") ~ "bn",
        TRUE ~ "bc"
      ),
      # If sample type is bn, remove b_n to save sample id only.
      sampleName = case_when(
        sampleType == "bn" ~ str_remove(sample, "_n$"),
        TRUE ~ sample
      ),
      sampleName = case_when(
        sampleName == "3302T1b_p" ~ "3302T1p",
        TRUE ~ sampleName
      ),
      # If sample type is bc, remove only _p, or _2 to match sample name edit of bn samples. Because we have b_p and b in some cases, we will need to loop through unique $sample and then extract the corresponding sample name edit for the matched bulk. 
      sampleNameEdit = case_when(
        sampleType == "bc" ~ str_remove(sampleName, "_p$|_2$"),
        TRUE ~ sampleName
      ),
       # BUT, if sample name is 3302T1b, the sampleNameEdit should be 3302T1bp, so that there is no duplicate sampleNameEdit for the error calculations. TODO: When doing the bulk cell vs bulk nuc, we will need to check that the two bulk cell errors for are grouped together. We could also duplicate the bn results for this sample. 

    ) %>%
  select(sampleName, sampleType, manual.log.anot.fine, cellProportion, sampleNameEdit) %>%
  # Method details
  mutate("deconvBenchResType" = "P", # P for proportion
         "Method" = "AutoGeneS", 
         "Parameters" = "fullCellTypeExpr", 
         "Reference" = "subSampled",
         "Seed"=123)

# Rbind proportion outputs to compute error metrics.
autogenesResGround <- rbind(autogenesResP, snCellPercentagesSubSampled)

# snCellPercentages have already been rbound to masterPropDf in the "MuSiC" chunk above, so will filter these out after error calculations.
masterPropDf <- rbind(masterPropDf, autogenesResP)


# # Compute and rbind metric outputs
# autogenesResMetrics <- deconError(autogenesResGround, 
#                               parameters = "fullCellTypeExpr",
#                               method = "AutoGeneS", 
#                               reference="subSampled",
#                               seed=123)
# # Rbind metric measurements 
# masterMetricDf <- rbind(masterMetricDf, autogenesResMetrics)


# Cleanup
rm(autogenesResGround, autogenesResP, autogenesResMetrics, AutoGeneS_full)
gc()
```

# 4 Save results
```{r}
write.xlsx(masterPropDf, paste0(resDir, fprefix, "_masterProportionDf.xlsx"))
# write.xlsx(masterMetricDf, paste0(resDir, fprefix, "_masterMetricDf.xlsx"))
```


# 5 Visualizations
## 5.1 Read in results 
```{r}
masterPropDf <- read.xlsx(paste0(resDir, fprefix, "_masterProportionDf.xlsx")) %>%
  mutate("Method_Param"=paste0(Method, "_", Parameters))

# masterMetricDf <- read.xlsx(paste0(resDir, fprefix, "_masterMetricDf.xlsx")) %>%
  mutate("Method_Param"=paste0(Method, "_", Parameters))
```

## 5.2 Global Metrics 
```{r}
# Of note, not all bulk samples were in snData
# > setdiff(dfPred$sampleNameEdit, dfTruth$sampleNameEdit)
# [1] "3177T4" "3177T5" "3177T6"
# > # Weirdly, saving with write.xlsx preserves rownames, but can't read them in? 
# > snCountsFinal_colData <- read.xlsx(paste0(dataDir, "01_Deconvolution_geneAnnotationsCTVTRefsnCountsFinal_colData.xlsx"))
# > bulkCountsFinal_metadata <- read.xlsx(paste0(dataDir, "01_Deconvolution_matchedBulkMatrixCTVTf_bulkCountsFinal_metadata.xlsx"))
# > View(bulkCountsFinal_metadata)
# > setdiff(bulkCountsFinal_metadata$sampleName, snCountsFinal_colData$Sample)
#  [1] "3176T3"    "3176T3b_n" "3177T2"    "3177T2a_n" "3177T4"    "3177T4a_n" "3177T5"    "3177T5a_n" "3177T6"    "3177T6a_n"
# [11] "3178T1"    "3178T1b_n" "3179T2"    "3179T2a_n" "3179T3"    "3179T3a_n" "3180T3"    "3180T3a_n" "3181T1"    "3181T1a_n"
# --> "3177T4"    "3177T4a_n"
# --> "3177T5"    "3177T5a_n"
# --> "3177T6"    "3177T6a_n"
```

### 5.2.1 RMSE
#### 5.2.1.1 Heatmap functions
Inspired by https://www.nature.com/articles/s41467-023-41385-5#Fig3. 
Takes average, should we take median? 
```{r}
rmseByCellTypeMean <- function(df, pred_type) {
  
  dfPred <- df %>%
    filter(sampleType %in% c(pred_type)) %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion, 
           Method_Param)
  
  dfTruth <- df %>%
    filter(sampleType == "sn") %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion) %>%
    dplyr::rename(truth = cellProportion)
  
  # 7 single-cell samples, 10 bulk cell and 10 bulk nuc samples. 
  intersect <- intersect(dfPred$sampleNameEdit, dfTruth$sampleNameEdit)
  
  dfPredIntersect <- dfPred %>%
    filter(sampleNameEdit %in% c(intersect))
  
  dfTruthIntersect <- dfTruth %>%
    filter(sampleNameEdit %in% c(intersect))

  # Join and compute RMSE
  rmseDf <- dfPredIntersect %>%
    left_join(dfTruthIntersect, 
              by = c("sampleNameEdit", "manual.log.anot.fine")) %>%
    group_by(manual.log.anot.fine, 
             Method_Param) %>%
    summarise(RMSE = sqrt(mean(((cellProportion - truth) * 100)^2)), 
              .groups = "drop")
  
  return(rmseDf)
}

rmseByCellTypeMedian <- function(df, pred_type) {
  
  dfPred <- df %>%
    filter(sampleType %in% c(pred_type)) %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion, 
           Method_Param)
  
  dfTruth <- df %>%
    filter(sampleType == "sn") %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion) %>%
    dplyr::rename(truth = cellProportion)
  
  # 7 single-cell samples, 10 bulk cell and 10 bulk nuc samples. 
  intersect <- intersect(dfPred$sampleNameEdit, dfTruth$sampleNameEdit)
  
  dfPredIntersect <- dfPred %>%
    filter(sampleNameEdit %in% c(intersect))
  
  dfTruthIntersect <- dfTruth %>%
    filter(sampleNameEdit %in% c(intersect))

  # Join and compute RMSE
  rmseDf <- dfPredIntersect %>%
    left_join(dfTruthIntersect, 
              by = c("sampleNameEdit", "manual.log.anot.fine")) %>%
    group_by(manual.log.anot.fine, 
             Method_Param) %>%
    summarise(RMSE = sqrt(median(((cellProportion - truth) * 100)^2)), 
              .groups = "drop") 
  
  return(rmseDf)
}

rmseBySampleMean <- function(df, pred_type) {
  
  dfPred <- df %>%
    filter(sampleType %in% c(pred_type)) %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion, 
           Method_Param)
  
  dfTruth <- df %>%
    filter(sampleType == "sn") %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion) %>%
    dplyr::rename(truth = cellProportion)
  
  # 7 single-cell samples, 10 bulk cell and 10 bulk nuc samples. 
  intersect <- intersect(dfPred$sampleNameEdit, dfTruth$sampleNameEdit)
  
  dfPredIntersect <- dfPred %>%
    filter(sampleNameEdit %in% c(intersect))
  
  dfTruthIntersect <- dfTruth %>%
    filter(sampleNameEdit %in% c(intersect))

  # Join and compute RMSE
  rmseDf <- dfPredIntersect %>%
    left_join(dfTruthIntersect, 
              by = c("sampleNameEdit", "manual.log.anot.fine")) %>%
    group_by(sampleNameEdit, 
             Method_Param) %>%
    summarise(RMSE = sqrt(mean(((cellProportion - truth) * 100)^2)), 
              .groups = "drop") 
  
  return(rmseDf)
}

rmseBySampleMedian <- function(df, pred_type) {
  
  dfPred <- df %>%
    filter(sampleType %in% c(pred_type)) %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion, 
           Method_Param)
  
  dfTruth <- df %>%
    filter(sampleType == "sn") %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion) %>%
    dplyr::rename(truth = cellProportion)
  
  # 7 single-cell samples, 10 bulk cell and 10 bulk nuc samples. 
  intersect <- intersect(dfPred$sampleNameEdit, dfTruth$sampleNameEdit)
  
  dfPredIntersect <- dfPred %>%
    filter(sampleNameEdit %in% c(intersect))
  
  dfTruthIntersect <- dfTruth %>%
    filter(sampleNameEdit %in% c(intersect))

  # Join and compute RMSE
  rmseDf <- dfPredIntersect %>%
    left_join(dfTruthIntersect, 
              by = c("sampleNameEdit", "manual.log.anot.fine")) %>%
    group_by(sampleNameEdit, 
             Method_Param) %>%
    summarise(RMSE = sqrt(median(((cellProportion - truth) * 100)^2)), 
              .groups = "drop") 
  
  return(rmseDf)
}

rmseCombinedCellNucSampleMean <- function(df, pred_type) {
  
  dfPred <- df %>%
    # filter(sampleType %in% c(pred_type)) %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion, 
           Method_Param) %>%
    mutate(sampleName)
  
  dfTruth <- df %>%
    # filter(sampleType == "sn") %>%
    select(sampleNameEdit, 
           sampleName, 
           manual.log.anot.fine, 
           cellProportion) %>%
    dplyr::rename(truth = cellProportion)
  
  # 7 single-cell samples, 10 bulk cell and 10 bulk nuc samples. 
  intersect <- intersect(dfPred$sampleNameEdit, dfTruth$sampleNameEdit)
  
  dfPredIntersect <- dfPred %>%
    filter(sampleNameEdit %in% c(intersect))
  
  dfTruthIntersect <- dfTruth %>%
    filter(sampleNameEdit %in% c(intersect))

  # Join and compute RMSE
  rmseDf <- dfPredIntersect %>%
    left_join(dfTruthIntersect, 
              by = c("sampleNameEdit", "manual.log.anot.fine")) %>%
    group_by(sampleNameEdit, 
             Method_Param) %>%
    summarise(RMSE = sqrt(mean(((cellProportion - truth) * 100)^2)), 
              .groups = "drop") 
  
  return(rmseDf)
}
```
#### 5.2.2.1 Plot and save results
##### 5.2.2.1.1 By cell type - mean
```{r}
allRmseList <- list()

# Plot
pdf(paste0(plotDir, "/heatmaps/", fprefix, "_meanRmse_byCellType.pdf"), 
      height = 6, width = 6)

for (ptype in c("bn", "bc")) {
    
    name <- ifelse(ptype=="bn", "bulkNuc", "bulkCell")
    rmse <- rmseByCellTypeMean(df = masterPropDf, 
                               pred_type = ptype)
    
    # Cast into matrix for heatmap
    rmseMat <- rmse %>%
      pivot_wider(names_from = Method_Param, 
                  values_from = RMSE) %>%
      column_to_rownames("manual.log.anot.fine") %>%
      as.matrix()
    
    pheatmap(rmseMat,
             main = paste0("RMSE - Matched Samples\n", name),
             color = colorRampPalette(c("white","pink", "red"))(100),
             breaks = seq(0, 100, length.out = 101), 
             angle_col = 45, fontsize_col = 5) 
    
    # Save for export
    allRmseList[[name]] <- rmse
  }

dev.off()
```



##### 5.2.1.1.1 By cell type - median
```{r}
allRmseList <- list()

# Plot
pdf(paste0(plotDir, "/heatmaps/", fprefix, "_medianRmse_byCellType.pdf"), 
      height = 6, width = 6)

for (ptype in c("bn", "bc")) {
    
    name <- ifelse(ptype=="bn", "bulkNuc", "bulkCell")
    rmse <- rmseByCellTypeMedian(df = masterPropDf, 
                               pred_type = ptype)
    
    # Cast into matrix for heatmap
    rmseMat <- rmse %>%
      pivot_wider(names_from = Method_Param, 
                  values_from = RMSE) %>%
      column_to_rownames("manual.log.anot.fine") %>%
      as.matrix()
    
    pheatmap(rmseMat,
             main = paste0("RMSE - Matched Samples\n", name),
             color = colorRampPalette(c("white","pink", "red"))(100),
             # breaks = seq(0, 100, length.out = 101), 
             angle_col = 90, fontsize_col = 6) 
    
    # Save for export
    allRmseList[[name]] <- rmse
  }

dev.off()
```

##### 5.2.1.1.1 By sample - mean
```{r}
allRmseList <- list()

# Plot
pdf(paste0(plotDir, "/heatmaps/", fprefix, "_meanRmse_bySampleNameEdit.pdf"), 
      height = 6, width = 6)

for (ptype in c("bn", "bc")) {
    
    name <- ifelse(ptype=="bn", "bulkNuc", "bulkCell")
    rmse <- rmseBySampleMean(df = masterPropDf, 
                               pred_type = ptype)
    
    # Cast into matrix for heatmap
    rmseMat <- rmse %>%
      pivot_wider(names_from = Method_Param, 
                  values_from = RMSE) %>%
      column_to_rownames("sampleNameEdit") %>%
      as.matrix()
    
    pheatmap(rmseMat,
             main = paste0("RMSE - Matched Samples\n", name),
             color = colorRampPalette(c("white","pink", "red"))(100),
             breaks = seq(0, 100, length.out = 101), 
             angle_col = 90, fontsize_col = 6) 
    
    # Save for export
    allRmseList[[name]] <- rmse
  }

dev.off()
```

##### 5.2.1.1.1 By sample - median
```{r}
allRmseList <- list()

# Plot
pdf(paste0(plotDir, "/heatmaps/", fprefix, "_medianRmse_bySampleNameEdit.pdf"), 
      height = 6, width = 6)

for (ptype in c("bn", "bc")) {
    
    name <- ifelse(ptype=="bn", "bulkNuc", "bulkCell")
    rmse <- rmseBySampleMedian(df = masterPropDf, 
                               pred_type = ptype)
    
    # Cast into matrix for heatmap
    rmseMat <- rmse %>%
      pivot_wider(names_from = Method_Param, 
                  values_from = RMSE) %>%
      column_to_rownames("sampleNameEdit") %>%
      as.matrix()
    
    pheatmap(rmseMat,
             main = paste0("RMSE - Matched Samples\n", name),
             color = colorRampPalette(c("white","pink", "red"))(100),
             breaks = seq(0, 100, length.out = 101), 
             angle_col = 90, fontsize_col = 6) 
    
    # Save for export
    allRmseList[[name]] <- rmse
  }

dev.off()
```
##### 5.2.2.1.5 All samples - mean
```{r}
# Compute RMSE using adjusted sample names for with sn matrix. Then left join with original sample names for each result. Then rbind all results and use original sample name as rows in pheatmap. 
rmseNucDf <- masterPropDf %>%
    filter(sampleType %in% c("bn")) %>%
    select(sampleNameEdit, 
           sampleName) %>%
  unique()


rmseNuc <- rmseBySampleMean(masterPropDf, 
                            pred_type = "bn") %>%
  left_join(rmseNucDf) %>%
  select(-sampleNameEdit)
  
rmseCellDf <- masterPropDf %>%
    filter(sampleType %in% c("bc")) %>%
    select(sampleNameEdit, 
           sampleName) %>%
  unique()

rmseCell <- rmseBySampleMean(masterPropDf, pred_type = "bc")%>%
  left_join(rmseCellDf) %>%
  select(-sampleNameEdit)

rmseAllSamples <- rbind(rmseNuc, rmseCell)

   

rmseMat <- rmseAllSamples %>%
      pivot_wider(names_from = Method_Param, 
                  values_from = RMSE) %>%
      column_to_rownames("sampleName") %>%
      as.matrix()


pdf(paste0(plotDir, "/heatmaps/", fprefix, "_meanRmse_bySampleName_allMatchedSamples.pdf"), 
      height = 6, width = 6)
pheatmap(rmseMat,
             main = paste0("RMSE - Matched Samples\n"),
             color = colorRampPalette(c("white","pink", "red"))(100),
             breaks = seq(0, 100, length.out = 101), 
             angle_col = 90, fontsize_col = 6) 
dev.off()
```

##### 5.2.2.1.5 All samples - median
```{r}
# Compute RMSE using adjusted sample names for with sn matrix. Then left join with original sample names for each result. Then rbind all results and use original sample name as rows in pheatmap. 
rmseNucDf <- masterPropDf %>%
    filter(sampleType %in% c("bn")) %>%
    select(sampleNameEdit, 
           sampleName) %>%
  unique()


rmseNuc <- rmseBySampleMedian(masterPropDf, 
                            pred_type = "bn") %>%
  left_join(rmseNucDf) %>%
  select(-sampleNameEdit)
  
rmseCellDf <- masterPropDf %>%
    filter(sampleType %in% c("bc")) %>%
    select(sampleNameEdit, 
           sampleName) %>%
  unique()

rmseCell <- rmseBySampleMedian(masterPropDf, pred_type = "bc")%>%
  left_join(rmseCellDf) %>%
  select(-sampleNameEdit)

rmseAllSamples <- rbind(rmseNuc, rmseCell)

   

rmseMat <- rmseAllSamples %>%
      pivot_wider(names_from = Method_Param, 
                  values_from = RMSE) %>%
      column_to_rownames("sampleName") %>%
      as.matrix()


pdf(paste0(plotDir, "/heatmaps/", fprefix, "_medianRmse_bySampleName_allMatchedSamples_noRange.pdf"), 
      height = 6, width = 6)
pheatmap(rmseMat,
             main = paste0("RMSE - Matched Samples\n"),
             color = colorRampPalette(c("white","pink", "red"))(100),
             # breaks = seq(0, 100, length.out = 101), 
             angle_col = 90, fontsize_col = 6) 
dev.off()

rmseRefDf <- masterPropDf %>%
    filter(sampleType %in% c("sn")) %>%
  filter(manual.log.anot.fine %in% c("CTVT_cells")) %>%
    select(
           sampleName, 
           cellProportion) %>%
  unique()

annot_df <- rmseRefDf %>%
  filter(sampleName %in% rownames(rmseMat)) %>%
  column_to_rownames("sampleName") %>%
  dplyr::rename("CTVT_snProportion"="cellProportion")

# Reorder to match matrix rows
annot_df <- annot_df[rownames(rmseMat), , drop = FALSE]

# Create the row annotation with a continuous color scale
row_ha <- rowAnnotation(
  CTVT_proportion = annot_df$CTVT_snProportion,
  col = list(CTVT_snProportion = colorRamp2(
    range(annot_df$CTVT_snProportion, na.rm = TRUE),
    c("white", "purple")
  )),
  annotation_name_side = "top"
)
# Define min, median, and max RMSE values
rmse_range <- range(rmseMat, na.rm = TRUE)
rmse_mid <- median(rmseMat, na.rm = TRUE)

# Define color function correctly
rmse_col_fun <- colorRamp2(
  c(rmse_range[1], rmse_mid, rmse_range[2]),
  c("white", "pink", "red")
  
  
)

ht <- Heatmap(
  rmseMat,
  name = "Median RMSE",
  col = rmse_col_fun,
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  left_annotation = row_ha,
  heatmap_legend_param = list(title = "RMSE")
)


pdf(paste0(plotDir, "/heatmaps/", fprefix, "_medianRmse_bySampleName_allMatchedSamples_noRange_withCTVTProp.pdf"), 
      height = 6, width = 6)

print(ht)

dev.off()

```

### 5.2.2 All samples all cell types
#### 5.2.2.1 Line plots
##### 5.2.2.1.1 Bulk nucleus vs bulk cell estimates 
```{r}
bulkComparison <- ggplot(masterMetricDf, aes(x = bulkCellEstimate, y = bulkNucEstimate, color = Method_Param)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ Metric, scales = "free") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  theme_minimal() +
  labs(
    title = "Comparison of Estimates by Metric",
    x = "Prediction on Bulk Cell Samples",
    y = "Prediction on Bulk Nucleus Samples"
  ) 

ggsave(paste0(plotDir,"line/", fprefix, "_masterMetricDf_bulkNuc_vs_bulkCell.pdf"),
       bulkComparison,
       height = 5, 
       width = 9)
```
##### 5.2.2.1.2 Dummy data
```{r}
# Create dummy limits per metric to control axis scaling
limits_df <- data.frame(
  Metric = c("MAE", "MAECorr", "SCorr", "CCorr"),
  x_min = c(0, 0, -1, -1),
  x_max = c(1, 1, 1, 1),
  y_min = c(0, 0, -1, -1),
  y_max = c(1, 1, 1, 1)
)

# Join dummy limits to your main data
plot_df <- masterMetricDf %>%
  left_join(limits_df, by = "Metric")

# Plot using geom_blank() with dummy limits to force correct axis ranges
blankPlot <- ggplot(plot_df, aes(x = bulkCellEstimate, y = bulkNucEstimate)) +
  geom_blank(aes(x = x_min, y = y_min)) +
  geom_blank(aes(x = x_max, y = y_max)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  facet_wrap(~ Metric, scales = "free") +
  theme_minimal() +
  labs(
    title = "",
    x = "Bulk Cell Estimate",
    y = "Bulk Nucleus Estimate"
  )

ggsave(paste0(plotDir,"line/", fprefix, "_masterMetricDf_blankData.pdf"),
       blankPlot,
       height = 5, 
       width = 7)
```
#### 5.2.2.2 Tile plots (MAE, MAECorr, SCorr, CCorr)

##### 5.2.2.2.1 Bulk Nuc
```{r}
# Reshape data to long format
plot_df <- masterMetricDf %>%
  pivot_longer(cols = c(bulkCellEstimate, bulkNucEstimate), 
               names_to = "EstimateType", values_to = "Value") %>%
  mutate(
    Metric = factor(Metric, levels = c("MAE", "MAECorr", "SCorr", "CCorr")),
    Method_Param = fct_rev(factor(Method_Param))  # Reverse for y-axis
  )

# plot_metric_tile <- function(data, estimate_type, metric_name, fill_low, fill_high, reverse = FALSE) {
#   
#   df <- data %>%
#     filter(EstimateType == estimate_type, Metric == metric_name)
#   
#   p <- ggplot(df, aes(x = Metric, y = Method_Param, fill = Value)) +
#     geom_tile(color = "white") +
#     scale_y_discrete(expand = expansion(mult = c(0, 0))) +
#     scale_x_discrete(expand = expansion(mult = c(0, 0))) +
#     (if (reverse) scale_fill_gradient(high = fill_low, low = fill_high, na.value = "white")
#      else scale_fill_gradient(low = fill_low, high = fill_high, na.value = "white")) +
#     theme_minimal(base_size = 14) +
#     labs(x = NULL, y = NULL, fill = metric_name) +
#     ggtitle(metric_name) +
#     theme(
#       axis.text.x = element_blank(),
#       axis.ticks.x = element_blank(),
#       strip.text = element_text(size = 14),
#       plot.title = element_text(hjust = 0.5),
#       legend.title = element_blank()
#     )
#   
#   return(p)
# }

plot_metric_tile <- function(data, 
                             estimate_type, 
                             metric_name, 
                             fill_low, 
                             fill_high, 
                             title,
                             reverse = FALSE) {
  
  df <- data %>%
    filter(EstimateType == estimate_type, Metric == metric_name)
  
  # Define the color scale limits based on the metric name
  color_limits <- if (metric_name %in% c("SCorr", "CCorr")) {
    c(-1, 1)  # for SCorr or CCorr
  } else {
    c(0, 1)   # for MAE or MAECorr
  }
  
  # Create the plot
  # p <- ggplot(df, aes(x = Metric, y = Method_Param_Reference, fill = Value)) +
  #   geom_tile(color = "white") +
  #   scale_y_discrete(expand = expansion(mult = c(0, 0))) +
  #   scale_x_discrete(
  #   expand = expansion(mult = c(0, 0)),
  #   labels = function(x) str_wrap(x, width = 10)
  # ) +
  #   (if (reverse) 
  #       scale_fill_gradient(low = fill_high, high = fill_low, na.value = "white", limits = color_limits)
  #    else 
  #       scale_fill_gradient(low = fill_low, high = fill_high, na.value = "white", limits = color_limits)) +
  #   theme_minimal(base_size = 14) +
  #   labs(x = NULL, y = NULL, fill = metric_name) +
  #   ggtitle(title) +
  #   theme(
  #     axis.text.x = element_blank(),
  #     axis.ticks.x = element_blank(),
  #     strip.text = element_text(size = 14),
  #     plot.title = element_text(hjust = 0.5),
  #     legend.title = element_blank(),
  #     axis.text.y = element_text(angle = 45, size = 8, hjust = 1)  # Rotate y-axis labels
  #   )
  
  # With text 
  p <- ggplot(df, aes(x = Metric, y = Method_Param, fill = Value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Value, 2)), size = 3) +  # <-- Add this line
  scale_y_discrete(expand = expansion(mult = c(0, 0))) +
  scale_x_discrete(
    expand = expansion(mult = c(0, 0)),
    labels = function(x) str_wrap(x, width = 10)
  ) +
  (if (reverse) 
      scale_fill_gradient(low = fill_high, high = fill_low, na.value = "white", limits = color_limits)
   else 
      scale_fill_gradient(low = fill_low, high = fill_high, na.value = "white", limits = color_limits)) +
  theme_minimal(base_size = 14) +
  labs(x = NULL, y = NULL, fill = metric_name) +
  ggtitle(title) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(size = 14),
    plot.title = element_text(hjust = 0.5),
    legend.title = element_blank(),
    axis.text.y = element_text(angle = 45, size = 8, hjust = 1)
  )

  
  return(p)
}



# Bulk Nucleus Plots
p1_nuc <- plot_metric_tile(plot_df, 
                           "bulkNucEstimate", 
                           "MAE", 
                           fill_low = "firebrick3", 
                           fill_high = "lightyellow", 
                           title = "Nuc",
                           reverse = T) # Lower MAE is better. 
p2_nuc <- plot_metric_tile(plot_df, 
                           "bulkNucEstimate", 
                           "MAECorr", 
                           "navyblue", 
                           "lightblue", 
                           title = "Nuc",
                           reverse = T) # Lower MAECorr is better. 
p3_nuc <- plot_metric_tile(plot_df, 
                           "bulkNucEstimate", 
                           "SCorr", 
                           "deeppink", 
                           "mistyrose",
                           title = "Nuc",
                           reverse = F) # Higher Corr is better
p4_nuc <- plot_metric_tile(plot_df, 
                           "bulkNucEstimate", 
                           "CCorr", 
                           
                           "darkgreen", 
                           "palegreen", 
                           title = "Nuc",
                           reverse = F)# Higher Corr is better

# Combine into one row
# nucleus_plot <- p1_nuc + p2_nuc + p3_nuc + p4_nuc + plot_layout(ncol = 4) +
#   plot_annotation(title = "Metric Heatmap for Bulk Nucleus Estimates")
 
# bulknucMetricsPerMethodParam <- wrap_plots(p1_nuc,
#                    p2_nuc,
#                    p3_nuc,
#                    p4_nuc,
#                    ncol=4,
#                    axis_titles="collect",
#                    axes="collect") +
#   plot_layout(guides = "collect") &
#   theme(legend.position = "bottom") 
# 
# 
# ggsave(paste0(plotDir, 
#               "tile/", 
#               fprefix, "_bulkNucMetricsPerMethodParam.pdf"), 
#        bulknucMetricsPerMethodParam,
#        height = 5, 
#        width= 10)
```
##### 5.2.2.2.2 Bulk Cell
```{r}
# Bulk Cellleus Plots
p1_cell <- plot_metric_tile(plot_df, 
                            "bulkCellEstimate", 
                            "MAE", 
                            "firebrick3", 
                            "lightyellow", 
                            title = "Cell",
                            reverse = T)
p2_cell <- plot_metric_tile(plot_df, 
                            "bulkCellEstimate", 
                            "MAECorr", 
                            "navyblue", 
                            "lightblue", 
                            title = "Cell",
                            reverse = T)
p3_cell <- plot_metric_tile(plot_df, 
                            "bulkCellEstimate", 
                            "SCorr", 
                            "deeppink", 
                            "mistyrose",
                            title = "Cell",
                            reverse = F)
p4_cell <- plot_metric_tile(plot_df, 
                            "bulkCellEstimate", 
                            "CCorr", 
                            "darkgreen", 
                            "palegreen", 
                            title = "Cell",
                            reverse = F)

# Combine into one row
# cellleus_plot <- p1_cell + p2_cell + p3_cell + p4_cell + plot_layout(ncol = 4) +
#   plot_annotation(title = "Metric Heatmap for Bulk Cellleus Estimates")
 
# bulkcellMetricsPerMethodParam <- wrap_plots(p1_cell,
#                    p2_cell,
#                    p3_cell,
#                    p4_cell,
#                    ncol=4,
#                    axis_titles="collect",
#                    axes="collect") +
#   plot_layout(guides = "collect") &
#   theme(legend.position = "bottom") &
#   theme(
#     legend.text = element_text(size = 8),        
#     legend.title = element_text(size = 9)         
#   ) +
#   theme(
#   axis.text.x = element_blank(),
#   axis.ticks.x = element_blank(),
#   strip.text = element_text(size = 14),
#   plot.title = element_text(hjust = 0.5),
#   legend.title = element_blank(),
#   axis.text.y = element_text(angle = 45, 
#                              size = 10, 
#                              hjust = 1) 
# )
# 
# 
# ggsave(paste0(plotDir, 
#               "tile/", 
#               fprefix, "_bulkCellMetricsPerMethodParam.pdf"), 
#        bulkcellMetricsPerMethodParam,
#        height = 5, 
#        width= 10)
```
##### 5.2.2.2.2 Bulk Nuc and Cell (Nuc Cell Nuc Cell)

```{r}
# # Combine side-by-side
# combinedPlot <- wrap_plots(
#   bulknucMetricsPerMethodParam, 
#   bulkcellMetricsPerMethodParam, nrow=2) +
#   plot_layout(guides = "collect") &
#   theme(legend.position = "bottom")

(bulknuccellMetricsPerMethodParam <- wrap_plots(p1_nuc,
                                            p1_cell,
                                            p2_nuc,
                                            p2_cell,
                                            p3_nuc,
                                            p3_cell,
                                            p4_nuc,
                                            p4_cell,
                                            ncol=8,
                                            nrow=1,
                                            axis_titles="collect",
                                            axes="collect") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom") &
  theme(
    legend.text = element_text(size = 8),        
    legend.title = element_text(size = 9)         
  ) +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  strip.text = element_text(size = 14),
  plot.title = element_text(hjust = 0.5),
  # legend.title = element_blank(),
  axis.text.y = element_text(angle = 45, 
                             size = 10, 
                             hjust = 1) 
))

ggsave(paste0(plotDir, 
              "tile/", 
              fprefix, "_bulkNucCellMetricsPerMethodParam_long.pdf"), 
       bulknuccellMetricsPerMethodParam,
       height = 8, 
       width= 12)
```


#### 5.2.2.2 Tile plots ("RMSE", "MAE", "Pearsonr", "RPE")

## 5.3 Granular Metrics
```{r}

```
# ---
# 6 New visualizations
## 6.1 Global parameters
```{r}
scriptDir <- paste0(wDir, "/12_Scripts/deconBenchmark/")
source(paste0(scriptDir, "/01_Deconvolution_deconBenchmark_visualizationFunctions_CTVT.R"))
source(paste0(scriptDir, "/01_Deconvolution_deconBenchmark_metricFunctions_CTVT.R"))

manualColors <- c(
  "Myeloid"              = "#85C0F9",  # Red
  "CAF.Endo"        = "maroon",  # Blue
  "Tcells" = "seagreen",  # Green
  "Bcells"     = "#A95AA1",  # Purple
  "Plasma"      = "lightpink",  # Orange
  "bad" = "darkgrey",
  "CTVT"           = "#0F2080"   # Brown
)

# plotDirHmPropDf <- paste0(plotDir, "/heatmaps/propDf/")
cellTypeOrder <- c("CTVT", "CAF.Endo", "Myeloid", 
                   "Bcells", "Plasma",  "Tcells", "bad")
sampleGroups <- list(
  c("Ground truth",  "Reference"),
  # c("sn_3302T1p", "bc_3302T1p"),
  c("sn_3302T1b", "bn_3302T1b", "bc_3302T1b", "bc_3302T1p"),
  # c("sn_3302T1p", "bn_3302T1b", "bc_3302T1p"),
  c("sn_3621T","bn_3621T", "bc_3621T"),
  c("sn_3203T1A2c", "bn_3203T1A2c", "bc_3203T1A2c"),
  c("sn_3145T1c","bc_3145T1c" , "bn_3145T1c"),
  c("sn_3145T11a", "bn_3145T11a", "bc_3145T11a" ),
  c("bc_3203T1A2b",  "bn_3203T1A2b"),
  c("bc_3776T"),
  c("bc_3775T" ),
  c("bc_3774T"),
  c("bc_3772T"),
  c("bc_3770T"),
  c("bc_3764T"),
  c("bc_3761T"),
  c("bc_3760T"),
  c("bc_3759T"),
  c("bc_3752T"))

# Flatten list to get sampleOrder (same as yours)
sampleOrder <- unlist(sampleGroups)

# Determine where to insert gaps: after each group
rowGaps <- unit(rep(5, length(sampleGroups) - 1), "pt")

# refProp <- read.xlsx(paste0(wDir, "13_Results/relativeProportionCellTypesPerReference.xlsx")) %>%
#   column_to_rownames("cellTypes") %>%
#   as.data.frame()

# Filter only "sn" samples for true proportions df. Never changes. 
masterPropDf <- read.xlsx(paste0(resDir, fprefix, "_masterProportionDf.xlsx")) %>%
  mutate("Method_Param"=paste0(Method, "_", Parameters))


snDf <- masterPropDf %>%
    filter(sampleType == "sn")%>%
    mutate(typeName = paste0(sampleType,"_", sampleNameEdit))
```

## 6.1 Proportion dataframes
### 6.1.1 Reference proportions
For each reference (ex. snCounts), we want to compute the average prop of cell types across all samples. 
```{r}
snCell <- rowSums(snCellCounts)

snCellProp <- snCell / sum(snCell) 

snCellSubSampleRelProp <- rowSums(snCellCountsSubSampled)

snCellPropSubSampleRelProp <- snCellSubSampleRelProp / sum(snCellSubSampleRelProp) 
```

### 6.1.2 Per Method_Param
Per combination of parameters, 
```{r}
print(unique(masterPropDf$Reference))

for (ref in unique(masterPropDf$Reference)){
  
  print(paste0("Running reference: ", ref))
  
  # Create heatmap directory for this reference. 
  plotDirHmPropDf <- paste0(plotDir, "/heatmaps/", ref, "/")
  plotDirBpPropDf <- paste0(plotDir, "/bar/", ref, "/")
  plotDirBarPropDf <- paste0(plotDir, "/bar/", ref, "/")
  # Will not overwrite any existing directories. 
  dir.create(plotDirHmPropDf, recursive = T)
  dir.create(plotDirBpPropDf,recursive = T)
  dir.create(plotDirBarPropDf,recursive = T)
  
  #########################
  # Get "All samples", "Reference", and "Training data" cell type proportions. 
  #########################
  
  # if ref == "snCounts", rbind the fullRef row twice (because snCounts=fullRef -- not an LOO)
  
  
  
  refDf <- NULL
  if (ref == "subSampled"){
    refDf <- rbind(snCellProp, snCellPropSubSampleRelProp) %>%
      as.data.frame() %>%
      select(cellTypeOrder)


    # Add rownames back!
    # colnames(refDf) <- cellTypeOrder
    rownames(refDf) <- c("Ground truth", "Reference")

    }else{

    refDf <- refProp %>%
      select(c(fullRef, ref)) %>%
      t()
  }
  
  

  # To match sampleGroup names and order. --> For LOO data.
  rownames(refDf) <- c("Ground truth", "Reference")
  
  # To match cell type order in heatmapMatrix.
  refDf <- refDf[,cellTypeOrder]
  
  #########################
  # Subset masterPropDf on reference for creating heatmapMatrix.
  #########################
  
  # Add column for heatmap rownames to match sampleGroups. Also add column matching seed and   parameters combination
  masterPropDfEdit <- masterPropDf %>%
    filter(Reference==ref) %>%
    mutate(typeName = paste0(sampleType,"_", sampleNameEdit))
  
  # Remainder of results df. 
  resDf <- masterPropDfEdit %>%
    filter(sampleType != "sn")

  

  for (mpr in unique(resDf$Method_Param)){
    print(mpr)
    
    # Filter sn counts. 
    resDfTruth <- masterPropDfEdit %>%
      filter(sampleType == "sn")
    
    # Filter mpr
    resDfPredict <- masterPropDfEdit %>%
      filter(Method_Param==mpr)
    
    # # Filter training data proportions for this Method_Param and add to refDf
    # tempTrainProp <- masterPropAvgParamSeed %>%
    #   filter(Method_Param == mpr) %>%
    #   # select(-c(Parameters, Method_Param)) %>%
    #   ungroup %>%
    #   select(colnames(refDf)) %>%
    #   # TODO: Check why adding grouping variables?
    #   as.matrix()
    # 
    # rownames(tempTrainProp) <- "Training data"
    
    tempRefDf <- rbind(refDf)
   
    
    # Rbind results and ground truth. 
    tempPlotDf <- rbind(resDfPredict, resDfTruth) %>%
      filter(typeName %in% sampleOrder)
    
    # Check to make sure all typeNames have all cell types in df
    all_counts_equal <- length(unique(table(tempPlotDf$typeName))) == 1
    
    if (all_counts_equal==FALSE) {
      cat(" typeName groups have unequal counts.\n")
      print(table(tempPlotDf$typeName))  # Optional: see where they differ
    }
  
    # Pivot to wide format
    heatmapMatrix <- tempPlotDf %>%
      select(typeName, manual.log.anot.fine, cellProportion) %>%
      pivot_wider(names_from = manual.log.anot.fine, values_from = cellProportion) %>%
      # Use mutated seq type _ sample name edit for new row names.
      column_to_rownames("typeName") %>%
      select(cellTypeOrder) %>%
      as.matrix() %>%
      rbind(tempRefDf)
    
    # Check to make sure all rownames are included in the ordering column of the samples in the heatmap. 
    diff <- setdiff(rownames(heatmapMatrix), sampleOrder)
    
    if (length(diff)!=0) {
      cat(" Missing sample names in sampleOrder")
      print(diff)
    }
    
    # Check to make sure all ordering columns are in rownames. 
    diff <- setdiff(sampleOrder, rownames(heatmapMatrix))
    if (length(diff)!=0) {
      cat(" Missing sampleOrder names in heatMapmatrix rownames.")
      print(diff)  
    }
    
    # Reorder for plotting
    heatmapMatrix <- heatmapMatrix[sampleOrder, cellTypeOrder]
    
    # Ensure that the color vector matches matrix columns
    cellTypes <- colnames(heatmapMatrix)
    stopifnot(all(cellTypes %in% names(manualColors)))
    
    # Create column-specific color functions: white  manual hue
    col_fun_list <- setNames(
      lapply(cellTypes, function(cell) {
        vals <- heatmapMatrix[, cell]
        colorRamp2(
          c(min(vals, na.rm = TRUE), max(vals, na.rm = TRUE)),
          c("white", manualColors[[cell]])
        )
      }),
      cellTypes
    )
    
    # Create a heatmap per cell type - printing values 
    heatmap_list_values <- lapply(cellTypes, function(cell) {
      Heatmap(heatmapMatrix[, cell, drop = FALSE],
              name = cell,
              col = col_fun_list[[cell]],
              cluster_rows = FALSE,
              cluster_columns = FALSE,
              show_row_names = TRUE,
              show_column_names = TRUE, 
              column_names_rot = 45, 
              row_gap = rowGaps,
              row_split = rep(seq_along(sampleGroups), lengths(sampleGroups)),
          cell_fun = function(j, i, x, y, width, height, fill) {
      grid.text(
        sprintf("%.2f", heatmapMatrix[i, cell]),
        x, y,
        gp = gpar(fontsize = 10, col = "black")
      )
      })
    })
    
    # Create a heatmap per cell type - Not printing values
    heatmap_list <- lapply(cellTypes, function(cell) {
      Heatmap(heatmapMatrix[, cell, drop = FALSE],
              name = cell,
              col = col_fun_list[[cell]],
              cluster_rows = FALSE,
              cluster_columns = FALSE,
              show_row_names = TRUE,
              show_column_names = TRUE, 
              column_names_rot = 45, 
              row_gap = rowGaps,
              row_split = rep(seq_along(sampleGroups), lengths(sampleGroups)))
    })
    
    # Combine and draw the full heatmap
    combinedHeatmap <- Reduce(`+`, heatmap_list_values)
    combinedHeatmapNoValues <- Reduce(`+`, heatmap_list)
    
    pdf(paste0(plotDirHmPropDf, fprefix,"_", mpr, "_proportionsHeatmap.pdf"), 
        height=10, 
        width = 10)
    draw(combinedHeatmap, 
         heatmap_legend_side = "top", 
         column_title = mpr, 
         column_title_side = "bottom",
         column_title_gp = gpar(fontsize = 8, fontface = "bold")) 
    draw(combinedHeatmapNoValues, 
         heatmap_legend_side = "top", 
         column_title = mpr, 
         column_title_side = "bottom",
         column_title_gp = gpar(fontsize = 8, fontface = "bold")) 
    dev.off()
    
    
    
    
    
    ####### Create proportion bars instead of cells. ##########
    # -------------------- Create grouped stacked barplot --------------------

    # Convert heatmapMatrix to long format
    barPlotDf <- as.data.frame(heatmapMatrix) %>%
      rownames_to_column("typeName") %>%
      pivot_longer(-typeName, names_to = "cellType", values_to = "proportion")
    
    # Create sampleGroup labels based on sampleGroups list
    sampleGroupMap <- do.call(rbind, lapply(seq_along(sampleGroups), function(i) {
      data.frame(
        typeName = sampleGroups[[i]],
        sampleGroup = paste0("Group_", i)
      )
    }))
    
   # Ensure sample order is respected
  barPlotDf$typeName <- factor(barPlotDf$typeName, levels = rev(sampleOrder))  # rev() makes top-to-bottom match sampleGroups
  
  # Plot
  ggBar <- ggplot(barPlotDf, aes(y = typeName, x = proportion, fill = cellType)) +
    geom_bar(stat = "identity", position = "stack", width = 0.8) +
    scale_fill_manual(values = manualColors) +
    labs(
      title = paste("Cell Type Proportions -", mpr),
      y = "Sample",
      x = "Proportion"
    ) +
    theme_bw(base_size = 6) +
    theme(
      axis.text.y = element_text(size = 8),
      panel.grid = element_blank()
    )


    
    # Save the plot
    ggsave(filename = paste0(plotDirBarPropDf, 
                             fprefix, 
                             "_",
                             mpr, 
                             "_barplotFaceted.pdf"),
           plot = ggBar, height = 5, width = 6)

  }
  
}
```
## 6.4 RMSE 
### 6.3.1 RMSE Heatmaps
```{r}
masterPropDfNew <- masterPropDf 
```

#### 6.3.1.1 All samples with all cell type proportions. 
```{r}
# Compute RMSE using adjusted sample names for with sn matrix. Then left join with original sample names for each result. Then rbind all results and use original sample name as rows in pheatmap. 
rmseNucDf <- masterPropDfNew %>%
    filter(sampleType %in% c("bn")) %>%
    select(sampleNameEdit, 
           sampleName) %>%
  unique()


rmseNuc <- rmseBySampleMedian(masterPropDfNew, 
                            pred_type = "bn", 
                            col="Method_Param") %>%
  filter(sampleNameEdit %in% rmseNucDf$sampleNameEdit) %>%
  left_join(rmseNucDf) %>%
  select(-sampleNameEdit)
  
rmseCellDf <- masterPropDfNew %>%
    filter(sampleType %in% c("bc")) %>%
    select(sampleNameEdit, 
           sampleName) %>%
  unique()

rmseCell <- rmseBySampleMedian(masterPropDfNew, 
                               pred_type = "bc", 
                               col = "Method_Param")%>%
  filter(sampleNameEdit %in% rmseCellDf$sampleNameEdit) %>%
  left_join(rmseCellDf) %>%
  select(-sampleNameEdit)

rmseAllSamples <- rbind(rmseNuc, rmseCell)

rmseMat <- rmseAllSamples %>%
      pivot_wider(names_from = "Method_Param", 
                  values_from = RMSE) %>%
      column_to_rownames("sampleName") %>%
      as.matrix() 

# Step 1: Calculate proportions of all cell types
all_celltype_props <- masterPropDf %>%
  filter(sampleType %in% c("sn")) %>%
  select(sampleName, manual.log.anot.fine, cellProportion) %>%
  unique() %>%
  pivot_wider(
    names_from = manual.log.anot.fine,
    values_from = cellProportion,
    values_fill = 0  # fill missing with 0
  ) %>% 
  select(-"CTVT_cells") # For color mapping

# Step 2: Match row order of rmseMat
all_celltype_props <- all_celltype_props %>%
  filter(sampleName %in% rownames(rmseMat)) %>%
  column_to_rownames("sampleName")

# Reorder to match matrix rows
all_celltype_props <- all_celltype_props[rownames(rmseMat), , drop = FALSE]

# Step 3: Create color scales for each cell type
celltype_colors <- lapply(all_celltype_props, function(x) {
  colorRamp2(
    range(x, na.rm = TRUE),
    c("white", scales::hue_pal()(1))  # assign different color per column later
  )
})

# Assign unique colors to each cell type (using hue palette), except CTVT
celltype_names <- colnames(all_celltype_props)[colnames(all_celltype_props)!="CTVT_cells"]

hues <- scales::hue_pal()(length(celltype_names))

celltype_colors <- setNames(
  lapply(seq_along(celltype_names), function(i) {
    colorRamp2(range(all_celltype_props[[i]], na.rm = TRUE), c("white", hues[i]))
  }),
  celltype_names
)

# Step 4: Create row annotation for all cell types - add CTVT from chunk above back in!
rmseRefDf <- masterPropDf %>%
    filter(sampleType %in% c("sn")) %>%
  filter(manual.log.anot.fine %in% c("CTVT_cells")) %>%
    select(
           sampleName, 
           cellProportion) %>%
  unique()

annot_df <- rmseRefDf %>%
  filter(sampleName %in% rownames(rmseMat)) %>%
  column_to_rownames("sampleName") %>%
  dplyr::rename("CTVT_snProportion"="cellProportion")

# Reorder to match matrix rows
annot_df <- annot_df[rownames(rmseMat), , drop = FALSE]

multi_celltype_annotation <- rowAnnotation(
  df = cbind(all_celltype_props, 
             CTVT_proportion = annot_df$CTVT_snProportion),
  col = c(celltype_colors, 
          CTVT_proportion = colorRamp2(
    range(annot_df$CTVT_snProportion, na.rm = TRUE),
    c("white", "purple"))),
  annotation_name_side = "top"
)

# Step 5: Combine with CTVT annotation if you want both
combined_row_ha <- multi_celltype_annotation  # replace with next line if you want both
# combined_row_ha <- row_ha + multi_celltype_annotation

# Define min, median, and max RMSE values
rmse_range <- range(rmseMat, na.rm = TRUE)
rmse_mid <- median(rmseMat, na.rm = TRUE)

# Define color function correctly
rmse_col_fun <- colorRamp2(
  c(rmse_range[1], rmse_mid, rmse_range[2]),
  c("white", "pink", "red"))


###################################
# Code column information
###################################
# Extract column information
paramDf <- data.frame(
  colname = colnames(rmseMat),
  stringsAsFactors = FALSE
)

pattern <- "snCounts_bal([^_]+)_thresh([0-9.]+)_ran([0-9.]+)_var([0-9.]+)_samp(\\d+)_cells(\\d+)_seed(\\d+)_lr([0-9eE.-]+)_steps(\\d+)_cran([0-9.]+)"


paramDf$method <- str_extract(paramDf$colname, "^(AdRoit|AutoGeneS|BayesPrism|DWLS|FARDEEP|LinDeconSeq|MIXTURE|MuSiC|RNA-Sieve)")

paramDf$reference <- str_remove(str_extract(paramDf$colname, "_(Default|fullCellTypeExpr|fullSignatureMatrix|subSampledSignatureMatrix)"), "_")


# Subset only the levels that are present in the data
# Define full color maps
method_colors <- c("True" = "forestgreen", 
                   "False" = "orange")
thresh_colors <- c("0" = "blue", 
                   # "thresh0" = "red",
                   # "thresh5"="purple",
                   "0.2"="pink"
                   )

  # Subset only the levels that are present in the data
bal_colors_present <- bal_colors[names(bal_colors) %in% unique(paramDf$bal)]
thresh_colors_present <- thresh_colors[names(thresh_colors) %in% unique(paramDf$thresh)]

ha_column = HeatmapAnnotation(
  method = paramDf$method,
  reference = paramDf$reference,
  col = list(
    method = structure(colorRampPalette(c("darkseagreen1", "darkseagreen4"))(length(unique(paramDf$method))),
                    names = sort(unique(paramDf$method))),
    reference = structure(colorRampPalette(c("blue", "whitesmoke", "firebrick2"))(length(unique(paramDf$reference))),
                     names = sort(unique(paramDf$reference)))
  ),
  annotation_name_side = "left",
  annotation_legend_param = list(title_gp = gpar(fontsize = 8))
)



###################################
# Clust columns and rows = T 
###################################

# Step 6: Add annotation to Heatmap
htClustRowCol <- Heatmap(
  rmseMat,
  name = "Median RMSE",
  col = rmse_col_fun,
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_column_names = F, 
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  bottom_annotation = ha_column,
  left_annotation = combined_row_ha,
  heatmap_legend_param = list(title = "RMSE"),
  cell_fun = function(j, i, x, y, width, height, fill){
    grid.text(sprintf("%.1f", rmseMat[i, j]), x, y, gp = gpar(fontsize = 6))}
)


###################################
# Clust rows but not columns 
###################################

# Step 6: Add annotation to Heatmap
htClustRow <- Heatmap(
  rmseMat,
  name = "Median RMSE",
  col = rmse_col_fun,
  cluster_rows = TRUE,
  cluster_columns = F,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  left_annotation = combined_row_ha,
  show_column_names = F, 
  bottom_annotation = ha_column,
  heatmap_legend_param = list(title = "RMSE"),
  cell_fun = function(j, i, x, y, width, height, fill){
    grid.text(sprintf("%.1f", rmseMat[i, j]), x, y, gp = gpar(fontsize = 6))}
)

###################################
# Clust neither rows nor columns
###################################

samples <- c(
  "3176T3b",
  "3176T3",
  "3177T2a",
  "3177T2",
  "3178T1b",
  "3178T1",
  "3179T2a",
  "3179T2",
  "3179T3a",
  "3179T3",
  "3180T3a",
  "3180T3",
  "3181T1a",
  "3181T1"
)

rmseMatManual <- rmseMat[samples,]

# Step 6: Add annotation to Heatmap
ht <- Heatmap(
  rmseMatManual,
  name = "Median RMSE",
  col = rmse_col_fun,
  cluster_rows = F,
  cluster_columns = F,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  left_annotation = combined_row_ha,
  show_column_names = F, 
  bottom_annotation = ha_column,
  heatmap_legend_param = list(title = "RMSE"),
  cell_fun = function(j, i, x, y, width, height, fill){
    grid.text(sprintf("%.1f", rmseMat[i, j]), x, y, gp = gpar(fontsize = 6))}
)

###################################
# Save
###################################

# Step 7: Output to PDF
pdf(paste0(plotDir, "/heatmaps/", fprefix, "_medianRmse_withAllCelltypeProps_paramSeed.pdf"), 
    height = 10, width = 20)

draw(
    htClustRowCol,
    heatmap_legend_side = "right",         # RMSE legend
    
    annotation_legend_side = "bottom",     # Annotation legend (e.g., column info)
    merge_legend = FALSE                   # Keep separate legends
)

draw(
    htClustRow,
    heatmap_legend_side = "right",         # RMSE legsend
    annotation_legend_side = "bottom",     # Annotation legend (e.g., column info)
    merge_legend = FALSE                   # Keep separate legends
)

draw(
    ht,
    heatmap_legend_side = "right",         # RMSE legsend
    annotation_legend_side = "bottom",     # Annotation legend (e.g., column info)
    merge_legend = FALSE                   # Keep separate legends
)


dev.off()
```

# 6 New visualizations
## 6.1 Global parameters
```{r}
source(paste0(scriptDir, "/01_Deconvolution_deconBenchmark_visualizationFunctions_CTVT.R"))
source(paste0(scriptDir, "/01_Deconvolution_deconBenchmark_metricFunctions_CTVT.R"))
manualColors <- c(
  "Myeloid"              = "skyblue",  # Red
  "CAF.Endo"        = "#E41A1C",  # Blue
  "Tcells" = "forestgreen",  # Green
  "Bcells"     = "magenta",  # Purple
  "Plasma"      = "orange",  # Orange
  "bad" = "darkgrey",
  "CTVT"           = "purple"   # Brown
)

# plotDirHmPropDf <- paste0(plotDir, "/heatmaps/propDf/")
cellTypeOrder <- c("CTVT", "CAF.Endo", "Myeloid", 
                   "Bcells", "Plasma",  "Tcells", "bad")

sampleGroups <- list(
  c("Ground truth",  "Reference", "Training data"),
  # c("sn_3302T1p", "bc_3302T1p"),
  c("sn_3302T1b", "bn_3302T1b", "bc_3302T1b", "bc_3302T1p"),
  # c("sn_3302T1p", "bn_3302T1b", "bc_3302T1p"),
  c("sn_3621T","bn_3621T", "bc_3621T"),
  c("sn_3203T1A2c", "bn_3203T1A2c", "bc_3203T1A2c"),
  c("sn_3145T1c","bc_3145T1c" , "bn_3145T1c"),
  c("sn_3145T11a", "bn_3145T11a", "bc_3145T11a" ),
  c("bc_3203T1A2b",  "bn_3203T1A2b"),
  c("bc_3776T"),
  c("bc_3775T" ),
  c("bc_3774T"),
  c("bc_3772T"),
  c("bc_3770T"),
  c("bc_3764T"),
  c("bc_3761T"),
  c("bc_3760T"),
  c("bc_3759T"),
  c("bc_3752T"))
  

# Flatten list to get sampleOrder (same as yours)
sampleOrder <- unlist(sampleGroups)

# Determine where to insert gaps: after each group
rowGaps <- unit(rep(5, length(sampleGroups) - 1), "pt")

# refProp <- read.xlsx(paste0(wDir, "13_Results/relativeProportionCellTypesPerReference.xlsx")) %>%
#   column_to_rownames("cellTypes") %>%
#   as.data.frame()

# # Filter only "sn" samples for true proportions df. Never changes. 
# snDf <- masterPropDf %>%
#     filter(sampleType == "sn")%>%
#     mutate(typeName = paste0(sampleType,"_", sampleNameEdit))
```
## 6.2 Read in results
```{r}
masterGlobMetricDf <- read.xlsx(paste0(resDir, fprefix, "_masterGlobalMetricDf.xlsx"))

masterGranularMetricDf <- read.xlsx(paste0(resDir,  fprefix, "_masterGranularMetricDf.xlsx"))

masterPropDf <- read.xlsx(paste0(resDir,  fprefix, "_masterProportionDf.xlsx"))

masterRawPEDf <- read.xlsx(paste0(resDir,  fprefix, "_masterRawPEDf.xlsx"))
```

## 6.1 Proportion dataframes
### 6.1.1 Reference proportions
For each reference (ex. snCounts), we want to compute the average prop of cell types across all samples. 
```{r}
snCell <- rowSums(snCellCountsSubSampled)

snCellProp <- snCell / sum(snCell) 
```

### 6.1.2 Per Param_Seed
Per combination of parameters, 
```{r}
print(unique(masterPropDf$Reference))

for (ref in unique(masterPropDf$Reference)){
  
  print(paste0("Running reference: ", ref))
  
  # Create heatmap directory for this reference. 
  plotDirHmPropDf <- paste0(plotDirScaden, "/heatmaps/", ref, "/")
  plotDirBpPropDf <- paste0(plotDirScaden, "/bar/", ref, "/")
  plotDirBarPropDf <- paste0(plotDirScaden, "/bar/", ref, "/")
  # Will not overwrite any existing directories. 
  dir.create(plotDirHmPropDf, recursive = T)
  dir.create(plotDirBpPropDf,recursive = T)
  dir.create(plotDirBarPropDf,recursive = T)
  
  #########################
  # Get "All samples", "Reference", and "Training data" cell type proportions. 
  #########################
  
  # if ref == "snCounts", rbind the fullRef row twice (because snCounts=fullRef -- not an LOO)
  
  
  
  refDf <- NULL
  if (ref == "snCounts"){
    refDf <- rbind(snCellProp, snCellProp) %>%
      as.data.frame() %>%
      select(cellTypeOrder)
      
    
    # Add rownames back! 
    # colnames(refDf) <- cellTypeOrder
    rownames(refDf) <- c("Ground truth", "snCounts")
  
    }else{
      
    refDf <- refProp %>%
      select(c(fullRef, ref)) %>%
      t() 
  }

  # To match sampleGroup names and order. --> For LOO data.
  rownames(refDf) <- c("Ground truth", "Reference")
  
  # To match cell type order in heatmapMatrix.
  refDf <- refDf[,cellTypeOrder]
  
  #########################
  # Subset masterPropDf on reference for creating heatmapMatrix.
  #########################
  
  # Add column for heatmap rownames to match sampleGroups. Also add column matching seed and   parameters combination
  masterPropDfEdit <- masterPropDf %>%
    filter(Reference==ref) %>%
    mutate(typeName = paste0(sampleType,"_", sampleNameEdit), 
           Param_Seed = paste0(Parameters,"_", Seed))
  
  # Remainder of results df. 
  resDf <- masterPropDfEdit %>%
    filter(sampleType != "sn")

  ###########################################
  ### Loop through Method_Param_Reference ###
  ### Multiple parameters possible per reference 
  ### Scripts are split by method so not really looping through methods rn. 
  ###########################################
  # a <- setdiff( unique(resDf$Param_Seed), unique(masterPropAvgParamSeed$Param_Seed))

  for (mpr in unique(resDf$Param_Seed)){
    print(mpr)
    
    # Filter sn counts. 
    resDfTruth <- masterPropDfEdit %>%
      filter(sampleType == "sn")
    
    # Filter mpr
    resDfPredict <- masterPropDfEdit %>%
      filter(Param_Seed==mpr)
    
    # Filter training data proportions for this Param_Seed and add to refDf
    tempTrainProp <- snCellProp
    
    rownames(tempTrainProp) <- "Training data"
    
    tempRefDf <- rbind(refDf, 
                       tempTrainProp)
   
    
    # Rbind results and ground truth. 
    tempPlotDf <- rbind(resDfPredict, resDfTruth) %>%
      filter(typeName %in% sampleOrder)
    
    # Check to make sure all typeNames have all cell types in df
    all_counts_equal <- length(unique(table(tempPlotDf$typeName))) == 1
    
    if (all_counts_equal==FALSE) {
      cat(" typeName groups have unequal counts.\n")
      print(table(tempPlotDf$typeName))  # Optional: see where they differ
    }
  
    # Pivot to wide format
    heatmapMatrix <- tempPlotDf %>%
      select(typeName, manual.log.anot.fine, cellProportion) %>%
      pivot_wider(names_from = manual.log.anot.fine, values_from = cellProportion) %>%
      # Use mutated seq type _ sample name edit for new row names.
      column_to_rownames("typeName") %>%
      select(cellTypeOrder) %>%
      as.matrix() %>%
      rbind(tempRefDf)
    
    # Check to make sure all rownames are included in the ordering column of the samples in the heatmap. 
    diff <- setdiff(rownames(heatmapMatrix), sampleOrder)
    
    if (length(diff)!=0) {
      cat(" Missing sample names in sampleOrder")
      print(diff)
    }
    
    # Check to make sure all ordering columns are in rownames. 
    diff <- setdiff(sampleOrder, rownames(heatmapMatrix))
    if (length(diff)!=0) {
      cat(" Missing sampleOrder names in heatMapmatrix rownames.")
      print(diff)  
    }
    
    # Reorder for plotting
    heatmapMatrix <- heatmapMatrix[sampleOrder, cellTypeOrder]
    
    # Ensure that the color vector matches matrix columns
    cellTypes <- colnames(heatmapMatrix)
    stopifnot(all(cellTypes %in% names(manualColors)))
    
    # Create column-specific color functions: white  manual hue
    col_fun_list <- setNames(
      lapply(cellTypes, function(cell) {
        vals <- heatmapMatrix[, cell]
        colorRamp2(
          c(min(vals, na.rm = TRUE), max(vals, na.rm = TRUE)),
          c("white", manualColors[[cell]])
        )
      }),
      cellTypes
    )
    
    # Create a heatmap per cell type - printing values 
    heatmap_list_values <- lapply(cellTypes, function(cell) {
      Heatmap(heatmapMatrix[, cell, drop = FALSE],
              name = cell,
              col = col_fun_list[[cell]],
              cluster_rows = FALSE,
              cluster_columns = FALSE,
              show_row_names = TRUE,
              show_column_names = TRUE, 
              column_names_rot = 45, 
              row_gap = rowGaps,
              row_split = rep(seq_along(sampleGroups), lengths(sampleGroups)),
          cell_fun = function(j, i, x, y, width, height, fill) {
      grid.text(
        sprintf("%.2f", heatmapMatrix[i, cell]),
        x, y,
        gp = gpar(fontsize = 10, col = "black")
      )
      })
    })
    
    # Create a heatmap per cell type - Not printing values
    heatmap_list <- lapply(cellTypes, function(cell) {
      Heatmap(heatmapMatrix[, cell, drop = FALSE],
              name = cell,
              col = col_fun_list[[cell]],
              cluster_rows = FALSE,
              cluster_columns = FALSE,
              show_row_names = TRUE,
              show_column_names = TRUE, 
              column_names_rot = 45, 
              row_gap = rowGaps,
              row_split = rep(seq_along(sampleGroups), lengths(sampleGroups)))
    })
    
    # Combine and draw the full heatmap
    combinedHeatmap <- Reduce(`+`, heatmap_list_values)
    combinedHeatmapNoValues <- Reduce(`+`, heatmap_list)
    
    pdf(paste0(plotDirHmPropDf, fprefix,"_", mpr, "_proportionsHeatmap.pdf"), 
        height=10, 
        width = 10)
    draw(combinedHeatmap, 
         heatmap_legend_side = "top", 
         column_title = mpr, 
         column_title_side = "bottom",
         column_title_gp = gpar(fontsize = 14, fontface = "bold")) 
    draw(combinedHeatmapNoValues, 
         heatmap_legend_side = "top", 
         column_title = mpr, 
         column_title_side = "bottom",
         column_title_gp = gpar(fontsize = 14, fontface = "bold")) 
    dev.off()
    
    
    
    
    
    ####### Create proportion bars instead of cells. ##########
    # -------------------- Create grouped stacked barplot --------------------

    # Convert heatmapMatrix to long format
    barPlotDf <- as.data.frame(heatmapMatrix) %>%
      rownames_to_column("typeName") %>%
      pivot_longer(-typeName, names_to = "cellType", values_to = "proportion")
    
    # Create sampleGroup labels based on sampleGroups list
    sampleGroupMap <- do.call(rbind, lapply(seq_along(sampleGroups), function(i) {
      data.frame(
        typeName = sampleGroups[[i]],
        sampleGroup = paste0("Group_", i)
      )
    }))
    
   # Ensure sample order is respected
  barPlotDf$typeName <- factor(barPlotDf$typeName, levels = rev(sampleOrder))  # rev() makes top-to-bottom match sampleGroups
  
  # Plot
  ggBar <- ggplot(barPlotDf, aes(y = typeName, x = proportion, fill = cellType)) +
    geom_bar(stat = "identity", position = "stack", width = 0.8) +
    scale_fill_manual(values = manualColors) +
    labs(
      title = paste("Cell Type Proportions -", mpr),
      y = "Sample",
      x = "Proportion"
    ) +
    theme_bw(base_size = 12) +
    theme(
      axis.text.y = element_text(size = 8),
      panel.grid = element_blank()
    )


    
    # Save the plot
    ggsave(filename = paste0(plotDirBarPropDf, 
                             fprefix, 
                             "_",
                             mpr, 
                             "_barplotFaceted.pdf"),
           plot = ggBar, height = 5, width = 6)

  }
  
}
```


## 6.4 RMSE 
### 6.3.1 RMSE Heatmaps
```{r}
masterPropDfNew <- masterPropDf 
```

#### 6.3.1.1 All samples with all cell type proportions. 
```{r}
# Compute RMSE using adjusted sample names for with sn matrix. Then left join with original sample names for each result. Then rbind all results and use original sample name as rows in pheatmap. 


rmseNucDf <- masterPropDfNew %>%
    filter(sampleType %in% c("bn")) %>%
    select(sampleNameEdit, 
           sampleName) %>%
  
  unique()


rmseNuc <- rmseBySampleMedian(masterPropDfNew, 
                            pred_type = "bn", 
                            col="Method_Param") %>%
  filter(sampleNameEdit %in% rmseNucDf$sampleNameEdit) %>%
  left_join(rmseNucDf) %>%
  select(-sampleNameEdit)%>%
  mutate(sampleName = paste0("bn_", sampleName))

  
rmseCellDf <- masterPropDfNew %>%
    filter(sampleType %in% c("bc")) %>%
    select(sampleNameEdit, 
           sampleName)%>%
  unique()

rmseCell <- rmseBySampleMedian(masterPropDfNew, 
                               pred_type = "bc", 
                               col = "Method_Param")%>%
  filter(sampleNameEdit %in% rmseCellDf$sampleNameEdit) %>%
  left_join(rmseCellDf) %>%
  select(-sampleNameEdit) %>%
  mutate(sampleName = paste0("bc_", sampleName))

rmseAllSamples <- rbind(rmseNuc, rmseCell)

rmseMat <- rmseAllSamples %>%
      pivot_wider(names_from = "Method_Param", 
                  values_from = RMSE) %>%
      column_to_rownames("sampleName") %>%
      as.matrix()

# Step 1: Calculate proportions of all cell types
all_celltype_props <- masterPropDf %>%
  filter(sampleType %in% c("sn")) %>%
  select(sampleNameEdit, manual.log.anot.fine, cellProportion) %>%
  unique() %>%
  pivot_wider(
    names_from = manual.log.anot.fine,
    values_from = cellProportion,
    values_fill = 0  # fill missing with 0
  ) 

# Step 2: Match row order of rmseMat
all_celltype_props <- all_celltype_props %>%
  filter(sampleName %in% rownames(rmseMat)) %>%
  column_to_rownames("sampleName")

# Reorder to match matrix rows
all_celltype_props <- all_celltype_props[colnames(rmseMat), , drop = FALSE]

# Step 3: Create color scales for each cell type
celltype_colors <- lapply(all_celltype_props, function(x) {
  colorRamp2(
    range(x, na.rm = TRUE),
    c("white", scales::hue_pal()(1))  # assign different color per column later
  )
})

# Assign unique colors to each cell type (using hue palette), except CTVT
celltype_names <- colnames(all_celltype_props)

hues <- scales::hue_pal()(length(celltype_names))

celltype_colors <- setNames(
  lapply(seq_along(celltype_names), function(i) {
    colorRamp2(range(all_celltype_props[[i]], na.rm = TRUE), c("white", hues[i]))
  }),
  celltype_names
)

# Step 4: Create row annotation for all cell types - add CTVT from chunk above back in!
rmseRefDf <- masterPropDf %>%
    filter(sampleType %in% c("sn")) %>%
  filter(manual.log.anot.fine %in% c("CTVT")) %>%
    select(
           sampleName, 
           cellProportion) %>%
  unique()

annot_df <- rmseRefDf %>%
  filter(sampleName %in% rownames(rmseMat)) %>%
  column_to_rownames("sampleName") %>%
  dplyr::rename("CTVT_snProportion"="cellProportion")

# Reorder to match matrix rows
annot_df <- annot_df[rownames(rmseMat), , drop = FALSE]

multi_celltype_annotation <- rowAnnotation(
  df = cbind(all_celltype_props),
  col = c(celltype_colors),
  annotation_name_side = "top"
)

# Step 5: Combine with CTVT annotation if you want both
combined_row_ha <- multi_celltype_annotation  # replace with next line if you want both
# combined_row_ha <- row_ha + multi_celltype_annotation

# Define min, median, and max RMSE values
rmse_range <- range(rmseMat, na.rm = TRUE)
rmse_mid <- median(rmseMat, na.rm = TRUE)

# Define color function correctly
rmse_col_fun <- colorRamp2(
  c(rmse_range[1], rmse_mid, rmse_range[2]),
  c("white", "pink", "red"))


###################################
# Code column information
###################################
# Extract column information
paramDf <- data.frame(
  colname = colnames(rmseMat),
  stringsAsFactors = FALSE
)

paramDf$method <- str_extract(paramDf$colname, "^(AdRoit|AutoGeneS|BayesPrism|DWLS|FARDEEP|LinDeconSeq|MIXTURE|MuSiC|RNA-Sieve)")

paramDf$reference <- str_remove(str_extract(paramDf$colname, "_(Default|fullCellTypeExpr|fullSignatureMatrix|subSampledSignatureMatrix)"), "_")


# Subset only the levels that are present in the data
# Define full color maps
method_colors <- c("True" = "forestgreen", 
                   "False" = "orange")
thresh_colors <- c("0" = "blue", 
                   # "thresh0" = "red",
                   # "thresh5"="purple",
                   "0.2"="pink"
                   )

  # Subset only the levels that are present in the data
bal_colors_present <- bal_colors[names(bal_colors) %in% unique(paramDf$bal)]
thresh_colors_present <- thresh_colors[names(thresh_colors) %in% unique(paramDf$thresh)]

ha_column = HeatmapAnnotation(
  method = paramDf$method,
  reference = paramDf$reference,
  col = list(
    method = structure(colorRampPalette(c("darkseagreen1", "darkseagreen4"))(length(unique(paramDf$method))),
                    names = sort(unique(paramDf$method))),
    reference = structure(colorRampPalette(c("blue", "whitesmoke", "firebrick2"))(length(unique(paramDf$reference))),
                     names = sort(unique(paramDf$reference)))
  ),
  annotation_name_side = "left",
  annotation_legend_param = list(title_gp = gpar(fontsize = 8))
)






###################################
# Clust columns and rows = T 
###################################

# Step 6: Add annotation to Heatmap
htClustRowCol <- Heatmap(
  rmseMat,
  name = "Median RMSE",
  col = rmse_col_fun,
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_column_names = F, 
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  bottom_annotation = ha_column,
  left_annotation = combined_row_ha,
  heatmap_legend_param = list(title = "RMSE"),
  cell_fun = function(j, i, x, y, width, height, fill){
    grid.text(sprintf("%.1f", rmseMat[i, j]), x, y, gp = gpar(fontsize = 6))}
)


###################################
# Clust rows but not columns 
###################################

# Step 6: Add annotation to Heatmap
htClustRow <- Heatmap(
  rmseMat,
  name = "Median RMSE",
  col = rmse_col_fun,
  cluster_rows = TRUE,
  cluster_columns = F,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  left_annotation = combined_row_ha,
  show_column_names = F, 
  bottom_annotation = ha_column,
  heatmap_legend_param = list(title = "RMSE"),
  cell_fun = function(j, i, x, y, width, height, fill){
    grid.text(sprintf("%.1f", rmseMat[i, j]), x, y, gp = gpar(fontsize = 6))}
)

###################################
# Clust neither rows nor columns
###################################

samples <- intersect(sampleOrder, rownames(rmseMat))

rmseMatManual <- rmseMat[samples,]

# Step 6: Add annotation to Heatmap
ht <- Heatmap(
  rmseMatManual,
  name = "Median RMSE",
  col = rmse_col_fun,
  cluster_rows = F,
  cluster_columns = F,
  row_names_gp = gpar(fontsize = 8),
  column_names_gp = gpar(fontsize = 8),
  left_annotation = combined_row_ha,
  show_column_names = F, 
  bottom_annotation = ha_column,
  heatmap_legend_param = list(title = "RMSE"),
  cell_fun = function(j, i, x, y, width, height, fill){
    grid.text(sprintf("%.1f", rmseMat[i, j]), x, y, gp = gpar(fontsize = 6))}
)

###################################
# Save
###################################

# Step 7: Output to PDF
pdf(paste0(plotDir, "/heatmaps/", fprefix, "_medianRmse_withAllCelltypeProps_paramSeed.pdf"), 
    height = 10, width = 20)

draw(
    htClustRowCol,
    heatmap_legend_side = "right",         # RMSE legend
    
    annotation_legend_side = "bottom",     # Annotation legend (e.g., column info)
    merge_legend = FALSE                   # Keep separate legends
)

draw(
    htClustRow,
    heatmap_legend_side = "right",         # RMSE legsend
    annotation_legend_side = "bottom",     # Annotation legend (e.g., column info)
    merge_legend = FALSE                   # Keep separate legends
)

draw(
    ht,
    heatmap_legend_side = "right",         # RMSE legsend
    annotation_legend_side = "bottom",     # Annotation legend (e.g., column info)
    merge_legend = FALSE                   # Keep separate legends
)


dev.off()
```


